## Spatial/OLS modeling function (by species)
See which species show significant spatial autocorrelation and use LeGrange test to estimate best spatial model to use

### Create spdf
```{r}
spatial.bumbles.glmvars.df
spatial.corr.test.df <- spatial.bumbles.glmvars.df %>%
  mutate(state_full = abbr2state(state)) %>%
  left_join(countycode.df,
            by = c("state_full" = "state",
                   "county" = "county_name"))
spatial.corr.test.spdf <- sp::merge(midwest.spdf,
                                  spatial.corr.test.df,
                                  by = "GID_2",
                                  duplicateGeoms = TRUE)

table(spatial.corr.test.spdf@data$species)

# Filter for study species in 2012 only
spatial.corr.test2012.spdf <- spatial.corr.test.spdf %>%
  filter(bin_ag == 2012) %>%
  filter(species %in% species)
```

### Find ag census year for each specices with most the records (counties)
```{r}
spatial.corr.test.df %>%
  group_by(species, bin_ag) %>%
  summarise(n_counties = n()) %>%
  filter(n_counties == max(n_counties),
         !species %in% c("Bombus ashtoni",
                         "Bombus fraternus",
                         "Bombus variabilis"))
```

### Test for spatial autocorrelation 
```{r}
spatial.autocorr.test <- function(spdf) {
  analysis.spdf <- spdf %>%
    filter(!is.na(species)) %>%
    filter(!is.na(prop_cropland))
  spp.test.yr.df <- spatial.corr.test.df %>%
    group_by(species, bin_ag) %>%
    summarise(n_counties = n()) %>%
    filter(n_counties == max(n_counties),
           !species %in% c("Bombus ashtoni",
                           "Bombus fraternus",
                           "Bombus variabilis"))
  species <- c("Bombus affinis",
               "Bombus auricomus",
               "Bombus bimaculatus",
               "Bombus borealis",
               "Bombus citrinus",
               "Bombus fervidus",
               "Bombus griseocollis",
               "Bombus impatiens",
               "Bombus pensylvanicus",
               "Bombus rufocinctus",
               "Bombus ternarius",
               "Bombus terricola",
               "Bombus vagans")
  iter <- 0
  for (i in species) {
    iter <- iter + 1
    bin_ag_max <- spp.test.yr.df$bin_ag[spp.test.yr.df$species == i]
    model.spdf <- analysis.spdf %>%
      filter(species == i) %>%
      filter(bin_ag == bin_ag_max)
    spdf.list[[iter]] <- model.spdf
  }
  names(spdf.list) <- species
  assign("spatial.autocorr.spdf.list",
         spdf.list,
         envir = .GlobalEnv)
  
  listnumber = 0
  modelresults.list <- list()
  modeloutput.list <- list()
  for (i in spdf.list) {
    listnumber <- listnumber + 1
    # OLS Model
    glm.model <- glm(rel_abun ~ n_crops + prop_cropland,
                     data = i@data,
                     weights = i@data$bin_ag_abundance,
                     na.action = na.exclude)
    glm.df <- tidy(glm.model) %>%
      mutate(model.type = "GLM",
             species = i@data$species[1],
             time = i@data$bin_ag[1],
             sample.size = nrow(i@data),
             moran.pvalue = NA) %>%
      dplyr::select(species, time, sample.size, model.type, moran.pvalue, everything())
    modeloutput.list[[listnumber]] <- glm.model
    # print(paste("n residuals:", length(resid(glm.model)), sep = " "))
    # Create spatial weights matrix
    neighbors <- poly2nb(i,
                         queen = TRUE)
    # print(paste("n neighbors", length(neighbors), sep = " "))
    weights <- nb2listw(neighbors,
                        # style = "W",
                        zero.policy = TRUE)
    moran.test <- lm.morantest(glm.model,
                               weights,
                               alternative = "two.sided",
                               zero.policy = TRUE)
    moran.pvalue <- moran.test$p.value[1, 1]
    glm.df$moran.pvalue = moran.pvalue
    modelresults.list[[listnumber]] <- glm.df
  }
  spatial.corr.df <- bind_rows(modelresults.list)
}

spatial.corr.test.df <- spatial.autocorr.test(spatial.corr.test.spdf)
```

## Spatial models function: 1870-2018
```{r}
spatial.corr.models <- function(spdf) {
  analysis.spdf <- spdf %>%
    filter(!is.na(species))
  spatial.test.list <- list()
  spdf.list <- list()
  species <- c("Bombus affinis",
               "Bombus griseocollis",
               "Bombus impatiens",
               "Bombus ternarius")
  for (i in species) {
    model.spdf <- analysis.spdf %>%
      filter(species == i) %>%
      filter(!is.na(prop_cropland))
    spdf.list[[i]] <- model.spdf
    # assign("test.list",
    #        spdf.list,
    #        envir = .GlobalEnv)
  }
  listnumber <- 0
  modelresults.list <- list()
  modeloutput.list <- list()
  sma.results.list <- list()
  sma.models.list <- list()
  for (i in spdf.list) {
    listnumber <- listnumber + 1
    glm.model <- glm(rel_abun ~ 
                       n_crops + 
                       prop_cropland + 
                       bin_ag,
                     data = i@data,
                     weights = i@data$bin_ag_abundance,
                     family = binomial(link = "logit"),
                     na.action = na.exclude)
    glm.df <- tidy(glm.model) %>%
      mutate(model.type = "glm",
             species = i@data$species[1],
             obsv = nrow(i@data)) %>%
      dplyr::select(species, model.type, obsv, everything())
    # residuals(ols.model)
    modeloutput.list[[listnumber]] <- glm.model
    weights.queen.list <- poly2nb(i,
                                  queen = TRUE)
    # print(length(weights.queen.list))
    # print(length(ols.model$residuals))
    # acf(residuals(ols.model))
    weights <- nb2listw(weights.queen.list,
                        style = "W",
                        zero.policy = TRUE)
    # plot(weights, coordinates(i))
    moran.test <- lm.morantest(glm.model,
                               weights,
                               alternative = "two.sided",
                               zero.policy = TRUE)
    moran.pvalue <- moran.test$p.value[1, 1]
    # print(moran.test)
    glm.df <- glm.df %>%
      mutate(moran_pvalue = moran.test$p.value)
    modelresults.list[[listnumber]] <- glm.df
    fit.sma <- errorsarlm(rel_abun ~ 
                            scale(n_crops) + 
                            scale(prop_cropland) + 
                            scale(bin_ag),
                          data = i,
                          weights = bin_ag_abundance,
                          listw = weights,
                          zero.policy = TRUE)
    sma.models.list[[listnumber]] <- fit.sma
    sma.model.df <- tibble(term = c("(Intercept)",
                                    "n_crops",
                                    "prop_cropland",
                                    "bin_ag"),
                           estimate = (summary(fit.sma)$Coef[, 1]),
                           std.error = (summary(fit.sma)$Coef[, 2]),
                           statistic = (summary(fit.sma)$Coef[, 3]),
                           p.value = (summary(fit.sma)$Coef[, 4])) %>%
      mutate(model.type = "spatial.error",
             species = i@data$species[1],
             sample.size = nrow(i@data)) %>%
      mutate(moran_pvalue = moran.test$p.value) %>%
      left_join(as_tibble(confint(fit.sma)) %>%
                  mutate(term = c("lamda",
                                  "(Intercept)",
                                  "n_crops",
                                  "prop_cropland",
                                  "bin_ag")) %>%
                  dplyr::select(term, lower.ci = `2.5 %`, upper.ci = `97.5 %`),
                by = c("term" = "term")) %>%
      dplyr::select(species, 
                    model.type, 
                    sample.size, 
                    moran_pvalue, 
                    term, 
                    estimate, 
                    lower.ci, 
                    upper.ci, 
                    z.stat = statistic,
                    p.value)
    print(sma.model.df)
    sma.results.list[[listnumber]] <- sma.model.df
  }
  glm.results.df <- bind_rows(modelresults.list)
  sma.results.df <- bind_rows(sma.results.list)
  assign("glm.spatialauto.df",
         glm.results.df,
         envir = .GlobalEnv)
  assign("sma.spatialauto.df",
         sma.results.df,
         envir = .GlobalEnv)
  assign("sma.models.list",
         sma.models.list,
         envir = .GlobalEnv)
}
```

### Run for 1870-2018
```{r}
spatial.corr.models(spatial.corr.test.spdf)
sma.spatialauto.df
glm.spatialauto.df
```

### Combine spatial models with GLMs
```{r}
# Coef's from glms.df already scaled.
models.1870.2018.df <- glms.df %>%
  filter(!species %in% c("Bombus affinis",
                         "Bombus griseocollis",
                         "Bombus impatiens",
                         "Bombus ternarius",
                         "Bombus ashtoni",
                         "Bombus fraternus",
                         "Bombus variabilis")) %>%
  mutate(model.type = "GLM",
         moran_pvalue = NA) %>%
  bind_rows(sma.spatialauto.df) %>%
  dplyr::select(species, 
                model.type, 
                sample.size, 
                moran_pvalue, 
                term, 
                estimate, 
                lower.ci, 
                upper.ci, 
                z.stat,
                p.value) %>%
  mutate(p.value = ifelse(p.value < 0.001,
                          "< 0.001",
                          round(p.value,
                                digits = 3)),
         estimate = round(estimate,
                          digits = 3),
         lower.ci = round(lower.ci,
                          digits = 4),
         upper.ci = round(upper.ci,
                          digits = 4))
models.1870.2018.df <- models.1870.2018.df %>%
  left_join(models.1870.2018.df %>%
              filter(term == "bin_ag") %>%
              mutate(status = case_when(estimate > 0 & p.value < 0.05 ~ "increase",
                                        estimate < 0 & p.value < 0.05 ~ "decrease",
                                        p.value > 0.05 ~ "stable"))%>%
              dplyr::select(species, status),
            by = "species") %>%
    write_csv("./model_output/sppglm_1870_2018.csv")
```
### Plot coefficients
```{r}
coef.1870.2018.plot <- models.1870.2018.df %>%
  filter(term != "(Intercept)") %>%
  group_by(term) %>%
  arrange(desc(estimate)) %>%
  mutate(row_order = factor(species, levels = species)) %>%
  ggplot() + 
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "gray60",
             size = 0.6) + 
  geom_pointrange(mapping = aes(x = row_order,
                                y = estimate,
                                ymin = lower.ci,
                                ymax = upper.ci),
                  color = "black",
                  size = 0.6) + 
  geom_point(mapping = aes(x = row_order,
                           y = estimate,
                           color = status),
             size = 8) +
  scale_color_manual(values = c("#E8C913", 
                                "#000000", 
                                "#E5E5E5"),
                     labels = c("Decreasing",
                                "Increasing",
                                "Stable"),
                     name = "Species status") +
  facet_wrap(~term) +
  theme_histbumbles() +
  theme(legend.position = "none") +
  coord_flip()
  ggsave("./model_plots/coefs/coef_1870_2018.eps",
       coef.1870.2018.plot,
       height = 4)
```

## Spatial models function: 1982-2018
### Create spdf
```{r}
spatial.corr.test.1982.df <- spatial.bumbles.glmvars2.df %>%
  mutate(state_full = abbr2state(state)) %>%
  left_join(countycode.df,
            by = c("state_full" = "state",
                   "county" = "county_name"))
spatial.corr.test.1982.spdf <- sp::merge(midwest.spdf,
                                  spatial.corr.test.1982.df,
                                  by = "GID_2",
                                  duplicateGeoms = TRUE)
```

```{r}
spatial.corr.1982.models <- function(spdf) {
  analysis.spdf <- spdf %>%
    filter(!is.na(species))
  spatial.test.list <- list()
  spdf.list <- list()
  species <- c("Bombus affinis",
               "Bombus griseocollis",
               "Bombus impatiens")
  for (i in species) {
    model.spdf <- analysis.spdf %>%
      filter(species == i) %>%
      filter(!is.na(prop_cc))
    spdf.list[[i]] <- model.spdf
    # assign("test.list",
    #        spdf.list,
    #        envir = .GlobalEnv)
  }
  listnumber <- 0
  modelresults.list <- list()
  modeloutput.list <- list()
  sma.results.list <- list()
  sma.models.list <- list()
  for (i in spdf.list) {
    listnumber <- listnumber + 1
    glm.model <- glm(rel_abun ~ 
                       n_crops + 
                       prop_pasture + 
                       prop_pest + 
                       prop_cc + 
                       bin_ag,
                     data = i@data,
                     weights = i@data$bin_ag_abundance,
                     family = binomial(link = "logit"),
                     na.action = na.exclude)
    glm.df <- tidy(glm.model) %>%
      mutate(model.type = "glm",
             species = i@data$species[1],
             obsv = nrow(i@data)) %>%
      dplyr::select(species, model.type, obsv, everything())
    modeloutput.list[[listnumber]] <- glm.model
    weights.queen.list <- poly2nb(i,
                                  queen = TRUE)
    # print(length(weights.queen.list))
    # print(length(ols.model$residuals))
    # acf(residuals(ols.model))
    weights <- nb2listw(weights.queen.list,
                        style = "W",
                        zero.policy = TRUE)
    # plot(weights, coordinates(i))
    moran.test <- lm.morantest(glm.model,
                               weights,
                               alternative = "two.sided",
                               zero.policy = TRUE)
    moran.pvalue <- moran.test$p.value[1, 1]
    # print(moran.test)
    glm.df <- glm.df %>%
      mutate(moran_pvalue = moran.test$p.value)
    modelresults.list[[listnumber]] <- glm.df
    fit.sma <- errorsarlm(rel_abun ~ 
                            scale(n_crops) + 
                            scale(prop_pasture) + 
                            scale(prop_pest) + 
                            scale(prop_cc) + 
                            scale(bin_ag),
                          data = i,
                          weights = bin_ag_abundance,
                          listw = weights,
                          zero.policy = TRUE)
    sma.models.list[[listnumber]] <- fit.sma
    sma.model.df <- tibble(term = c("(Intercept)",
                                    "n_crops",
                                    "prop_pasture",
                                    "prop_pest",
                                    "prop_cc",
                                    "bin_ag"),
                           estimate = (summary(fit.sma)$Coef[, 1]),
                           std.error = (summary(fit.sma)$Coef[, 2]),
                           statistic = (summary(fit.sma)$Coef[, 3]),
                           p.value = (summary(fit.sma)$Coef[, 4])) %>%
      mutate(model.type = "spatial.error",
             species = i@data$species[1],
             sample.size = nrow(i@data)) %>%
      mutate(moran_pvalue = moran.test$p.value) %>%
      left_join(as_tibble(confint(fit.sma)) %>%
                  mutate(term = c("lamda",
                                  "(Intercept)",
                                  "n_crops",
                                  "prop_pasture",
                                  "prop_pest",
                                  "prop_cc",
                                  "bin_ag")) %>%
                  dplyr::select(term, lower.ci = `2.5 %`, upper.ci = `97.5 %`),
                by = c("term" = "term")) %>%
      dplyr::select(species, 
                    model.type, 
                    sample.size, 
                    moran_pvalue, 
                    term, 
                    estimate, 
                    lower.ci, 
                    upper.ci, 
                    z.stat = statistic,
                    p.value)
    print(sma.model.df)
    sma.results.list[[listnumber]] <- sma.model.df
  }
  glm.results.df <- bind_rows(modelresults.list)
  sma.results.df <- bind_rows(sma.results.list)
  assign("glm.spatialauto.1982.df",
         glm.results.df,
         envir = .GlobalEnv)
  assign("sma.spatialauto.1982.df",
         sma.results.df,
         envir = .GlobalEnv)
  assign("sma.models.1982.list",
         sma.models.list,
         envir = .GlobalEnv)
}
```

### Run for 1982-2018
```{r}
spatial.corr.1982.models(spatial.corr.test.1982.spdf)
sma.spatialauto.1982.df
```

### Combine spatial models with GLMs
```{r}
# Coef's from glms.df already scaled.
models.1982.2018.df <- glms2.df %>%
  filter(!species %in% c("Bombus affinis",
                         "Bombus griseocollis",
                         "Bombus impatiens",
                         "Bombus ashtoni",
                         "Bombus fraternus",
                         "Bombus variabilis")) %>%
  mutate(model.type = "GLM",
         moran_pvalue = NA) %>%
  bind_rows(sma.spatialauto.1982.df %>%
              filter(species != "Bombus ternarius")) %>%
  dplyr::select(species, 
                model.type, 
                sample.size, 
                moran_pvalue, 
                term, 
                estimate, 
                lower.ci, 
                upper.ci, 
                z.stat,
                p.value) %>%
  mutate(p.value = ifelse(p.value < 0.001,
                          "< 0.001",
                          round(p.value,
                                digits = 3)),
         estimate = round(estimate,
                          digits = 3),
         lower.ci = round(lower.ci,
                          digits = 4),
         upper.ci = round(upper.ci,
                          digits = 4))

models.1982.2018.df <- models.1982.2018.df %>%
  left_join(models.1982.2018.df %>%
              filter(term == "bin_ag") %>%
              mutate(status = case_when(estimate > 0 & p.value < 0.05 ~ "increase",
                                        estimate < 0 & p.value < 0.05 ~ "decrease",
                                        p.value > 0.05 ~ "stable"))%>%
              dplyr::select(species, status),
            by = "species") %>%
    write_csv("./model_output/sppglm_1982_2018.csv")
```

### Plot coefficients
```{r}
coef.1982.2018.plot <- models.1982.2018.df %>%
  filter(term != "(Intercept)") %>%
  group_by(term) %>%
  arrange(desc(estimate)) %>%
  mutate(row_order = factor(species, levels = species)) %>%
  ggplot() + 
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "gray60",
             size = 0.6) + 
  geom_pointrange(mapping = aes(x = row_order,
                                y = estimate,
                                ymin = lower.ci,
                                ymax = upper.ci),
                  color = "black",
                  size = 0.6) + 
  geom_point(mapping = aes(x = row_order,
                           y = estimate,
                           color = status),
             size = 8) +
  scale_color_manual(values = c("#E8C913", 
                                "#000000", 
                                "#E5E5E5"),
                     labels = c("Decreasing",
                                "Increasing",
                                "Stable"),
                     name = "Species status") +
  # scale_y_continuous(limits = c(-2, 10),
  #                    breaks = seq(-2, 10, by = 1)) +
  facet_wrap(~term,
             nrow = 1,
             scales = "free_x") +
  theme_histbumbles() +
  theme(legend.position = "none") +
  coord_flip()
ggsave("./model_plots/coefs/coef_1982_2018.eps",
       coef.1982.2018.plot,
       height = 4,
       width = 10)
```

