---
title: "Ch4 | Historical Ag & Bumble Bee Abundance/Community Changes"
subtitle: "Historical bumble bee records data cleanup & prep"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_notebook
---

**Name:** Jeremy Hemberger
**Email:** j.hemberger.wisc@gmail.com
**Institution:** University of Wisconsin - Madison | Department of Entomology

# Import packages
```{r} 
library(tidyverse)
library(directlabels)
library(rgdal)
library(ggmap)
library(ggalt)
library(raster)
library(maptools)
library(maps)
library(mapdata)
library(mp)
library(vegan)
library(magrittr)
library(openintro)
```

# Clean up bumble bee data
## Load in raw data
```{r}
gbif.bumbles.raw <- read_tsv("./2018_GBIF_Bumbles.csv",
                             col_names = TRUE,
                            na = "NA")
xerc.bumbles.raw <- read_csv("./BBW_US_5Sept2018.csv",
                             col_names = TRUE,
                             na = "NA")
irc.bumbles.raw <- read_csv("./2016_WIRC_Bombus.csv",
                            col_names = TRUE, 
                            na = "NA") 
```

Glance at data for cleanup
```{r}
glimpse(gbif.bumbles.raw)
glimpse(xerc.bumbles.raw)
glimpse(irc.bumbles.raw)
```


## Standardize data structure: 

```{r}
gbif.bumbles <-
  tibble(
  unique.id = seq(from = 1, to = nrow(gbif.bumbles.raw)),
  og_database = "gbif",
  database_id_1 = gbif.bumbles.raw$gbifid,
  database_id_2 = gbif.bumbles.raw$occurrenceid,
  database_id_3 = gbif.bumbles.raw$catalognumber,
  collector = gbif.bumbles.raw$recordedby,
  phylum = gbif.bumbles.raw$phylum,
  class = gbif.bumbles.raw$class,
  order = gbif.bumbles.raw$order,
  family = gbif.bumbles.raw$family,
  genus = gbif.bumbles.raw$genus,
  species = gbif.bumbles.raw$species,
  id_to_taxon = gbif.bumbles.raw$taxonrank,
  sci_name = gbif.bumbles.raw$scientificname,
  country = gbif.bumbles.raw$countrycode,
  state = NA,
  county = NA,
  locality = gbif.bumbles.raw$locality,
  dec_lat = gbif.bumbles.raw$decimallatitude,
  dec_long = gbif.bumbles.raw$decimallongitude,
  elevation = gbif.bumbles.raw$elevation,
  date = gbif.bumbles.raw$eventdate,
  day = gbif.bumbles.raw$day,
  month = gbif.bumbles.raw$month,
  year  = gbif.bumbles.raw$year,
  institution_code = gbif.bumbles.raw$institutioncode,
  database_notes_1 = gbif.bumbles.raw$issue,
  database_notes_2 = NA,
  database_notes_3 = NA
  )
```

```{r}
xerc.bumbles <-
  tibble(
  unique.id = seq(from = 1, to = nrow(xerc.bumbles.raw)),
  og_database = "bb_watch",
  database_id_1 = xerc.bumbles.raw$`Bee Id`,
  database_id_2 = NA,
  database_id_3 = NA,
  collector = xerc.bumbles.raw$Observer,
  phylum = "Arthropoda",
  class =  "Insecta",
  order = "Hymenoptera",
  family = "Apidae",
  genus = "Bombus",
  species = xerc.bumbles.raw$Species,
  id_to_taxon = ifelse(is.na(xerc.bumbles.raw$Species), 
                       "GENUS", 
                       "SPECIES"),
  sci_name = paste(xerc.bumbles.raw$Genus, 
                   xerc.bumbles.raw$Species, 
                   sep = " "),
  country = xerc.bumbles.raw$Country,
  state = xerc.bumbles.raw$`Provinces/States`,
  county = NA,
  locality = xerc.bumbles.raw$Locality,
  dec_lat = xerc.bumbles.raw$Latitude,
  dec_long = xerc.bumbles.raw$Longitude,
  elevation = NA,
  date = as.character(xerc.bumbles.raw$`Date Observed`),
  day = NA,
  month = NA,
  year = NA,
  institution_code = xerc.bumbles.raw$Project, 
  database_notes_1 = xerc.bumbles.raw$`Floral Host`,
  database_notes_2 = xerc.bumbles.raw$Status,
  database_notes_3 = xerc.bumbles.raw$Private
  )
```

```{r}
irc.bumbles <-
  tibble(
  unique.id = seq(from = 1, to = nrow(irc.bumbles.raw)),
  og_database = "uw_irc",
  database_id_1 = paste("irc", 
                        nrow(irc.bumbles.raw),
                        sep = "."),
  database_id_2 = NA,
  database_id_3 = NA,
  collector = irc.bumbles.raw$Collector,
  phylum = "Arthropoda",
  class =  "Insecta",
  order = "Hymenoptera",
  family = "Apidae",
  genus = "Bombus",
  species = irc.bumbles.raw$Species,
  id_to_taxon = ifelse(is.na(irc.bumbles.raw$Species), 
                       "GENUS", 
                       "SPECIES"),
  sci_name = irc.bumbles.raw$Species,
  country = "USA",
  state = irc.bumbles.raw$State,
  county = irc.bumbles.raw$County,
  locality = irc.bumbles.raw$`City/Town`,
  dec_lat = NA,
  dec_long = NA,
  elevation = NA,
  date = irc.bumbles.raw$`Collection Date`, 
  day = NA,
  month = NA,
  year = NA,
  institution_code = "UW_WIRC",
  database_notes_1 = irc.bumbles.raw$Notes,
  database_notes_2 = irc.bumbles.raw$`Specific location (description)`,
  database_notes_3 = NA
  )

irc.bumbles$species <- str_replace(irc.bumbles$species,
                                   pattern = " \\(.*\\)",
                                   replacement = "")
irc.bumbles$species[irc.bumbles$species == "Bombus"] <- "Bombus auricomus"
irc.bumbles$species[irc.bumbles$species == "Bombus impatiens Cresson"] <- "Bombus impatiens"
irc.bumbles$species[irc.bumbles$species == "Bombus affinis Cresson"] <- "Bombus affinis"
```

Remove original raw data from environment
```{r}
rm(gbif.bumbles.raw)
rm(xerc.bumbles.raw)
rm(irc.bumbles.raw)
```

## Reformat data types
Dates to dates
IRC Dates need to be manually run through as they're in a fucking state...
```{r}
gbif.bumbles$date <- as.Date(gbif.bumbles$date, "%Y/%m/%d")
xerc.bumbles$date <- as.Date(xerc.bumbles$date, "%m/%d/%Y")
irc.bumbles$date <- as.Date(irc.bumbles$date, "%m/%d/%Y") 
```

Extract m/d/y to columns
```{r}
gbif.bumbles$day <- format(as.Date(gbif.bumbles$date), "%d")
gbif.bumbles$month <- format(as.Date(gbif.bumbles$date), "%m")
gbif.bumbles$year <- format(as.Date(gbif.bumbles$date), "%Y")

xerc.bumbles$day <- format(as.Date(xerc.bumbles$date), "%d")
xerc.bumbles$month <- format(as.Date(xerc.bumbles$date), "%m")
xerc.bumbles$year <- format(as.Date(xerc.bumbles$date), "%Y")

irc.bumbles$day <- format(as.Date(irc.bumbles$date), "%d")
irc.bumbles$month <- format(as.Date(irc.bumbles$date), "%m")
irc.bumbles$year <- format(as.Date(irc.bumbles$date), "%Y")
```

Clean up species names
```{r}
unique(gbif.bumbles$species)
unique(xerc.bumbles$species)
unique(irc.bumbles$species)
```

```{r}
xerc.bumbles$species <- paste(xerc.bumbles$genus,
                              xerc.bumbles$species,
                              sep = " ")
```

```{r}
xerc.bumbles$id_to_taxon <- ifelse(grepl("sp.",
                                         xerc.bumbles$species),
                                   "GENUS",
                                   "SPECIES")
```

# Filter out incomplete data
## Need lat/long
```{r}
ngbif.before <- nrow(gbif.bumbles)
gbif.bumbles <- gbif.bumbles %>%
  filter(!is.na(dec_lat) & !is.na(dec_long)) %>%
  filter(dec_long < 0 & dec_long > -174)
ngbif.after <- nrow(gbif.bumbles)
ngbif.before - ngbif.after
```

## Need ID to species
```{r}
ngbif.before <- nrow(gbif.bumbles)
gbif.bumbles <- gbif.bumbles %>%
  filter(id_to_taxon == "SPECIES")
ngbif.after <- nrow(gbif.bumbles)
ngbif.before - ngbif.after
```

```{r}
nxerc.before <- nrow(xerc.bumbles)
xerc.bumbles <- xerc.bumbles %>%
  filter(!is.na(dec_lat) & !is.na(dec_long)) %>%
  filter(state != "Alaska")
nxerc.after <- nrow(xerc.bumbles)
nxerc.before - nxerc.after
```
All BBWatch records have lat/long

```{r}
nxerc.before <- nrow(xerc.bumbles)
xerc.bumbles <- xerc.bumbles %>%
  filter(id_to_taxon == "SPECIES")
nxerc.after <- nrow(xerc.bumbles)
nxerc.before - nxerc.after
```
From 12,267 records to 11,275.  Diff of 992 records.

## Lower 48 only
```{r}
census.code <- read_csv("./national_county_code.txt",
                        col_names = c("state",
                                      "statefp",
                                      "countyfp",
                                      "countyname",
                                      "classfp"))
state.abbs <- tibble(abbreviation = state.abb, 
                     name = state.name,
                     region = state.region)
gbif.latlong <- tibble(gbif.bumbles$dec_long,
                       gbif.bumbles$dec_lat)
gbif.bumble.points <- SpatialPoints(gbif.latlong)

us.counties <- readOGR("./USCounties_Shapefile/", 
                       "cb_2015_us_county_5m")
proj4string(gbif.bumble.points) <- proj4string(us.counties)
gbif.bumbles$county <- over(gbif.bumble.points, us.counties)$NAME
gbif.bumbles$state <- over(gbif.bumble.points, us.counties)$STATEFP
gbif.bumbles$state <- census.code$state[match(as.vector(gbif.bumbles$state),
                                              census.code$statefp)]


gbif.bumbles <- gbif.bumbles %>%
  filter(state != "AK" & state != "HI")

plot(gbif.bumble.points)
```

Determine county for BBWatch
```{r}
xerc.latlong <- tibble(xerc.bumbles$dec_long,
                       xerc.bumbles$dec_lat)
xerc.bumble.points <- SpatialPoints(xerc.latlong)
proj4string(xerc.bumble.points) <- proj4string(us.counties)
xerc.bumbles$county <- over(xerc.bumble.points, us.counties)$NAME
```


Replace state names in bumble bee watch with state abbreviations
```{r}
xerc.bumbles$state <- state.abbs$abbreviation[match(as.vector(xerc.bumbles$state),
                                                    state.abbs$name)]
```


## Combine GBIF and BBWatch records to single data frame
```{r}
all.bumbles <- rbind(gbif.bumbles, xerc.bumbles)
all.bumbles$unique.id <- seq(from = 1,
                             to = nrow(all.bumbles))
all.bumbles$region <- state.abbs$region[match(as.vector(all.bumbles$state),
                                               state.abbs$abbreviation)]
write_csv(all.bumbles,
          "all_bumbles.csv")
```
Total records: 181,156

# Basic data summaries 
## Summary by state (n, n.species)
```{r}
all.bumbles %>%
  group_by(state) %>%
  summarise(n.species = n_distinct(species))

all.bumbles %>% 
  group_by(state) %>%
  summarise(n.bumbles = n())
```

## Summary by county (n, n.species)
```{r}
all.bumbles %>%
  group_by(state, county) %>%
  summarise(n.species = n_distinct(species)) %>%
  ggplot() +
  geom_histogram(mapping = aes(n.species)) + 
  theme_minimal()

all.bumbles %>%
  group_by(state, county) %>%
  summarise(n.bumbles = log(n())) %>%
  ggplot() +
  geom_histogram(mapping = aes(n.bumbles),
                 binwidth = 0.5) + 
  theme_minimal()

all.bumbles %>%
  group_by(state, county, species) %>%
  summarise(n.bumbles = n())
```


## Temporal record trends (n.decade)
```{r}
all.bumbles$day <- as.numeric(all.bumbles$day)
all.bumbles$month <- as.numeric(all.bumbles$month)
all.bumbles$year <- as.numeric(all.bumbles$year)
all.bumbles <- all.bumbles %>%
  ungroup() %>%
  mutate(decade = year - (year %% 10))
```
```{r}
all.bumbles %>%
  group_by(decade) %>%
  summarise(n.bumbles = n())

all.bumbles %>%
  group_by(og_database, decade) %>%
  summarise(n.bumbles = n()) %>%
  filter(n.bumbles > 500) %>%
  ggplot(mapping = aes(x = decade,
                       y = n.bumbles,
                       fill = og_database)) + 
  geom_col() + 
  theme_minimal()
```
```{r}
all.bumbles %>%
  group_by(state, decade) %>%
  summarise(n.bumbles = n())

all.bumbles %>%
  group_by(state, decade) %>%
  filter(decade > 1890) %>%
  filter(state %in% c("MN", # Filter to upper midwest/great lakes states
                      "WI", 
                      "IL",
                      "IA",
                      "ND",
                      "SD",
                      "MI",
                      "IN")) %>%
  summarise(n.bumbles = n()) %>%
  ggplot() + 
  geom_line(mapping = aes(x = decade,
                          y = n.bumbles,
                          color = state)) +
  theme_minimal()
```

## Create historical/contemporary periods and summarize
```{r}
all.bumbles <- all.bumbles %>%
  mutate(t_period = ifelse(decade <= 1999,
                           "historical",
                           "contemp"))
table(all.bumbles$t_period)
```

```{r}
hist.latlong <- all.bumbles %>%
  filter(t_period == "historical") %>%
  dplyr::select(dec_long,
                dec_lat)
hist.bumble.points <- SpatialPoints(hist.latlong)
contemp.latlong <- all.bumbles %>%
  filter(t_period == "contemp") %>%
  dplyr::select(dec_long,
                dec_lat)
contemp.bumble.points <- SpatialPoints(contemp.latlong)
plot(hist.bumble.points, pch = 1, col = "red")
plot(contemp.bumble.points, pch = 1, col = "blue", add = TRUE)
```

```{r}
bumbles.sum.df <- all.bumbles %>%
  ungroup() %>%
  group_by(county, state, t_period) %>%
  summarise(n.bumbles = n()) %>%
  filter(!is.na(t_period)) %>%
  filter(!is.na(county)) %>%
  mutate(state_name = abbr2state(state))
national.bumbles.df <- full_join(ungroup(bumbles.sum.df),
                                ungroup(us.county.48),
                                by = c("county" = "CNTY_NAME")) %>%
  replace_na(n.bumbles = 0)
national.bumbles.df <- national.bumbles.df %>%
  filter(!is.na(state))
national.bumbles.df$t_period <- factor(national.bumbles.df$t_period,
                                      levels = c("historical", "contemp"))
period.names <- c("historical" = "Historical Bumble Bee Records",
                  "contemp" = "Contemporary Bumble Bee Records")
national.bumbles.plot <- ggplot() + 
  geom_polygon(data = national.bumbles.df,
               mapping = aes(x = long,
                             y = lat,
                             fill = log(n.bumbles),
                             group = group)) +
  geom_polygon(data = us.county.48,
               mapping = aes(x = long,
                             y = lat,
                             group = group),
               fill = "transparent",
               color = "gray50",
               size = 0.05,
               na.rm = TRUE) + 
  scale_fill_distiller(palette = "Reds",
                       direction = 1,
                       na.value = "transparent") +
  coord_map("stereographic") +
  facet_grid(cols = vars(t_period),
             labeller = as_labeller(period.names)) + 
  theme_void()
national.bumbles.plot
```


```{r}
analysis.bumbles.df <- all.bumbles %>%
  filter(state %in% c("MN", # Filter to upper midwest/great lakes states
                      "WI", 
                      "IL",
                      "MI",
                      "IN", 
                      "IA"))
table(analysis.bumbles.df$t_period)
table(analysis.bumbles.df$species)
```

## Filter to relevant species, create 4 category temporal term
```{r}
analysis.bumbles.df <- analysis.bumbles.df %>%
  filter(species %in% c("Bombus auricomus",
                        "Bombus bimaculatus",
                        "Bombus fraternus",
                        "Bombus griseocollis",
                        "Bombus impatiens",
                        "Bombus pensylvanicus",
                        "Bombus vagans",
                        "Bombus variabilis",
                        "Bombus ternarius",
                        "Bombus affinis",
                        "Bombus terricola",
                        "Bombus borealis",
                        "Bombus ashtoni",
                        "Bombus fervidus",
                        "Bombus citrinus",
                        "Bombus rufocinctus")) %>%
  mutate(t_period = ifelse(year <= 1999,
                           "historical",
                           "contemp")) %>%
  mutate(time_period = ifelse(year <= 1930, 
                              "1900-1930",
                              ifelse(year <= 1960,
                                     "1931-1960",
                                     ifelse(year <= 1999,
                                            "1961-1999",
                                            "2000-present"))))
table(analysis.bumbles.df$time_period)
write_csv(analysis.bumbles.df,
          "./analysis_bumbles.csv")
analysis.bumbles.df
```

## Plot midwest n.bumbles by county
```{r}
analysis.bumbles.sum.df <- analysis.bumbles.df %>%
  ungroup() %>%
  group_by(county, state, time_period) %>%
  summarise(n.bumbles = n()) %>%
  filter(!is.na(time_period)) %>%
  filter(!is.na(county)) %>%
  mutate(state_name = abbr2state(state))
us.county.midwest <- us.county.48 %>%
  filter(region %in% c("Minnesota",
                       "Wisconsin",
                       "Illinois",
                       "Iowa",
                       "Michigan",
                       "Indiana"))
midwest.bumbles.df <- full_join(ungroup(analysis.bumbles.sum.df),
                                ungroup(us.county.midwest),
                                by = c("county" = "CNTY_NAME")) %>%
  replace_na(n.bumbles = 0)
```


```{r}
midwest.bumbles.df <- midwest.bumbles.df %>%
  filter(!is.na(state))
midwest.bumbles.df$time_period <- factor(midwest.bumbles.df$time_period,
                                      levels = c("1900-1930", 
                                                 "1931-1960",
                                                 "1961-1999",
                                                 "2000-present"))
period.names <- c("1900-1930" = "1900-1930", 
                  "1931-1960" = "1931-1960",
                  "1961-1999" = "1961-1999",
                  "2000-present" = "2000-present")
midwest.bumbles.plot <- ggplot() + 
  geom_polygon(data = midwest.bumbles.df,
               mapping = aes(x = long,
                             y = lat,
                             fill = log(n.bumbles),
                             group = group)) +
  geom_polygon(data = us.county.midwest,
               mapping = aes(x = long,
                             y = lat,
                             group = group),
               fill = "transparent",
               color = "gray80",
               size = 0.25,
               na.rm = TRUE) + 
  scale_fill_distiller(palette = "Reds",
                       direction = 1,
                       na.value = "transparent") +
  coord_map("stereographic") +
  facet_grid(cols = vars(time_period),
             labeller = as_labeller(period.names)) + 
  theme_void()
midwest.bumbles.plot
```

## Plot singleton counties by time period
```{r}
midwest.bumbles.single.df <- midwest.bumbles.df %>%
  mutate(singleton = ifelse(n.bumbles == 1,
                            1,
                            0))
midwest.singletons.plot <- ggplot() + 
  geom_polygon(data = midwest.bumbles.single.df,
               mapping = aes(x = long,
                             y = lat,
                             fill = factor(singleton),
                             group = group)) +
  geom_polygon(data = us.county.midwest,
               mapping = aes(x = long,
                             y = lat,
                             group = group),
               fill = "transparent",
               color = "gray80",
               size = 0.25,
               na.rm = TRUE) + 
  scale_fill_manual(values = c("1" = "red", "0" = "white")) +
  coord_map("stereographic") +
  facet_grid(cols = vars(time_period),
             labeller = as_labeller(period.names)) + 
  theme(legend.position = "none") + 
  theme_void()
midwest.singletons.plot
```

## Temporal trends in abundance by species
```{r}
analysis.bumbles.spsum.df <- analysis.bumbles.df %>%
  ungroup() %>%
  group_by(county, state, species, time_period) %>%
  summarise(n.bumbles = n()) %>%
  filter(!is.na(time_period)) %>%
  filter(!is.na(county)) %>%
  mutate(state_name = abbr2state(state)) %>%
  filter(species %in% c("Bombus auricomus",
                        "Bombus bimaculatus",
                        "Bombus fraternus",
                        "Bombus griseocollis",
                        "Bombus impatiens",
                        "Bombus pensylvanicus",
                        "Bombus vagans",
                        "Bombus variabilis",
                        "Bombus ternarius",
                        "Bombus affinis",
                        "Bombus terricola",
                        "Bombus borealis",
                        "Bombus ashtoni",
                        "Bombus fervidus",
                        "Bombus citrinus",
                        "Bombus rufocinctus"))
midwest.bumbles.spsum.df <- full_join(ungroup(analysis.bumbles.spsum.df),
                                      ungroup(us.county.midwest),
                                      by = c("county" = "CNTY_NAME")) %>%
  replace_na(n.bumbles = 0)
midwest.bumbles.spsum.df <- midwest.bumbles.spsum.df %>%
  filter(!is.na(state))
midwest.bumbles.df$time_period <- factor(midwest.bumbles.df$time_period,
                                      levels = c("1900-1930", 
                                                 "1931-1960",
                                                 "1961-1999",
                                                 "2000-present"))
midwest.bumbles.spsum.plot <- ggplot() + 
  geom_polygon(data = midwest.bumbles.spsum.df,
               mapping = aes(x = long,
                             y = lat,
                             fill = log(n.bumbles),
                             group = group)) +
  geom_polygon(data = us.county.midwest,
               mapping = aes(x = long,
                             y = lat,
                             group = group),
               fill = "transparent",
               color = "gray80",
               size = 0.25,
               na.rm = TRUE) + 
  scale_fill_distiller(palette = "Reds",
                       direction = 1,
                       na.value = "transparent") +
  coord_map("stereographic") +
  facet_grid(vars(time_period), 
             vars(species)) + 
  theme_void()
midwest.bumbles.spsum.plot
```


# Calculate diversity by county
```{r}
analysis.bumbles.df %>%
  ungroup() %>%
  group_by(county, state, species, t_period) %>%
  summarise(n.bumbles = n()) %>%
  filter(!is.na(t_period)) %>%
  filter(!is.na(county)) %>%
  mutate(state_name = abbr2state(state)) %>%
  filter(species %in% c("Bombus auricomus",
                        "Bombus bimaculatus",
                        "Bombus fraternus",
                        "Bombus griseocollis",
                        "Bombus impatiens",
                        "Bombus pensylvanicus",
                        "Bombus vagans",
                        "Bombus variabilis",
                        "Bombus ternarius",
                        "Bombus affinis",
                        "Bombus terricola",
                        "Bombus borealis",
                        "Bombus ashtoni",
                        "Bombus fervidus",
                        "Bombus citrinus",
                        "Bombus rufocinctus")) %>%
  spread(species,
         n.bumbles,
         fill = 0) %>%
  ungroup() %>%
  mutate(shannon = diversity(.[, 5:ncol(.)],
                             index = "shannon"),
         simpson = diversity(.[, 5:ncol(.)],
                             index = "simpson")) %>%
  # (county, t_period) %>%
  # summarise(mean_shannon = mean(shannon)) %>%
  dplyr::select(state_name, county, t_period, shannon) %>%
  spread(t_period, shannon) %>%
  filter(!is.na(contemp)) %>%
  filter(!is.na(historical)) %>%
  filter(historical != 0,
         contemp != 0) %>%
  mutate(diff = contemp - historical) %>%
  mutate(direction = ifelse(diff < 0,
                            "lost.diversity",
                            "gained.diversity")) %>%
  group_by(state_name) %>%
  arrange(desc(diff)) %>%
  ungroup() %>%
  mutate(state.county = paste(state_name, 
                              county, 
                              sep = "_")) %>%
  mutate(row_order = factor(.$`state.county`,
                            levels = .$`state.county`)) %>%
  ggplot() + 
  geom_segment(mapping = aes(x = row_order,
                             y = 0,
                             yend = diff,
                             xend = row_order),
               color = "black",
               alpha = 0.75) + 
  geom_point(mapping = aes(x = row_order,
                           y = diff,
                           col = direction),
             size = 3) + 
  scale_color_manual(name = "Diversity Trend",
                     labels = c("Gained Diversity", 
                                "Lost Diversity"),
                     values = c("lost.diversity" = "tomato2",
                                "gained.diversity" = "#0e668b")) + 
  theme_minimal() + 
  coord_flip() +
  facet_grid(vars(state_name))
```

# Plot change in relative abundance over time
## Calculate relative abundance by time period
```{r}
spp.abun.by.period.df <- analysis.bumbles.spsum.df %>%
  ungroup() %>%
  group_by(time_period, species) %>%
  summarise(total_of_spp = sum(n.bumbles))
spp.abun.by.period.df

relative.abun.df <- spp.abun.by.period.df %>%
  left_join(total.abun.by.period.df,
            by = "time_period") %>%
  mutate(relative_abun = total_of_spp / total_bumbles) %>%
  mutate(subgenus = ifelse(species %in% c("Bombus affinis",
                                          "Bombus terricola"),
                           "Bombus",
                           ifelse(species %in% c("Bombus griseocollis",
                                                 "Bombus rufocinctus",
                                                 "Bombus fraternus"),
                                  "Cullumanobombus",
                                  ifelse(species %in% c("Bombus vagans",
                                                        "Bombus bimaculatus",
                                                        "Bombus impatiens",
                                                        "Bombus ternarius",
                                                        "Bombus perplexus",
                                                        "Bombus sandersoni"),
                                         "Pyrobombus",
                                         ifelse(species %in% c("Bombus auricomus"),
                                                "Bombias",
                                                ifelse(species %in% c("Bombus fervidus",
                                                                      "Bombus pensylvanicus"),
                                                       "Thoracobombus",
                                                       ifelse(species %in% c("Bombus citrinus",
                                                                              "Bombus variabilis",
                                                                              "Bombus ashtoni"),
                                                              "Psithyrus",
                                                              "Subterraneobombus")))))))
relative.abun.df$time_period <- factor(relative.abun.df$time_period,
                                       levels = c("1900-1930", 
                                                  "1931-1960",
                                                  "1961-1999",
                                                  "2000-present"))
relative.abun.df
```

## Plot relative abundance changes
```{r}
relative.abun.df %>%
  ggplot() + 
  geom_point(mapping = aes(x = time_period,
                           y = relative_abun,
                           col = species),
             size = 3) + 
  # geom_line(mapping = aes(x = time_period,
  #                         y = relative_abun,
  #                         group = species),
  #           color = "#EEC643") + 
  geom_smooth(mapping = aes(x = time_period,
                            y = relative_abun,
                            col = species,
                            group = species),
              method = "lm",
              se = FALSE,
              color = "black",
              linetype = 1) + 
  theme_light() + 
  scale_y_continuous(limits = c(-0.05, 0.4)) + 
  scale_x_discrete(expand = c(0.05, 0.05)) + 
  coord_fixed(ratio = 2) + 
  facet_grid(vars(subgenus)) + 
  ggsave("relativeabun4.eps")
  

```

## Two-period relative abundance changes
```{r}
spsum.2period.df <- analysis.bumbles.df %>%
  ungroup() %>%
  group_by(county, state, species, t_period) %>%
  summarise(n.bumbles = n()) %>%
  filter(!is.na(t_period)) %>%
  filter(!is.na(county)) %>%
  mutate(state_name = abbr2state(state)) %>%
  filter(species %in% c("Bombus auricomus",
                        "Bombus bimaculatus",
                        "Bombus fraternus",
                        "Bombus griseocollis",
                        "Bombus impatiens",
                        "Bombus pensylvanicus",
                        "Bombus vagans",
                        "Bombus variabilis",
                        "Bombus ternarius",
                        "Bombus affinis",
                        "Bombus terricola",
                        "Bombus borealis",
                        "Bombus ashtoni",
                        "Bombus fervidus",
                        "Bombus citrinus",
                        "Bombus rufocinctus"))

total.abun.by.2period.df <- spsum.2period.df %>%
  ungroup() %>%
  group_by(t_period) %>%
  summarise(total_bumbles = sum(n.bumbles))
total.abun.by.2period.df

spp.abun.by.2period.df <- spsum.2period.df %>%
  ungroup() %>%
  group_by(t_period, species) %>%
  summarise(total_of_spp = sum(n.bumbles))
spp.abun.by.2period.df

relative.abun.2period.df <- spp.abun.by.2period.df %>%
  left_join(total.abun.by.2period.df,
            by = "t_period") %>%
  mutate(relative_abun = total_of_spp / total_bumbles) %>%
  mutate(subgenus = ifelse(species %in% c("Bombus affinis",
                                          "Bombus terricola"),
                           "Bombus",
                           ifelse(species %in% c("Bombus griseocollis",
                                                 "Bombus rufocinctus",
                                                 "Bombus fraternus"),
                                  "Cullumanobombus",
                                  ifelse(species %in% c("Bombus vagans",
                                                        "Bombus bimaculatus",
                                                        "Bombus impatiens",
                                                        "Bombus ternarius",
                                                        "Bombus perplexus",
                                                        "Bombus sandersoni"),
                                         "Pyrobombus",
                                         ifelse(species %in% c("Bombus auricomus"),
                                                "Bombias",
                                                ifelse(species %in% c("Bombus fervidus",
                                                                      "Bombus pensylvanicus"),
                                                       "Thoracobombus",
                                                       ifelse(species %in% c("Bombus citrinus",
                                                                              "Bombus variabilis",
                                                                              "Bombus ashtoni"),
                                                              "Psithyrus",
                                                              "Subterraneobombus")))))))
relative.abun.2period.df
relative.abun.2period.df$t_period <- factor(relative.abun.2period.df$t_period,
                                       levels = c("historical",
                                                  "contemp"))
relative.abun.2period.df %>%
  ggplot(mapping = aes(label = species)) + 
  geom_line(mapping = aes(x = t_period,
                          y = relative_abun,
                          color = species,
                          group = species),
            # color = "#EEC643",
            size = 1.25) +
  geom_point(mapping = aes(x = t_period,
                           y = relative_abun,
                           color = species),
             size = 3) + 
             # col = "#EEC643",
             # alpha = 0.75) + 
  # geom_text(mapping = aes(x = t_period,
  #                         y = relative_abun + 0.1),
  #           color = "black") + 
  # geom_dl(mapping = aes(x = t_period,
  #                       y = relative_abun,
  #                       label = species),
  #         method = "last.points",
  #         color = "black") + 
  # geom_smooth(mapping = aes(x = t_period,
  #                           y = relative_abun,
  #                           group = species),
  #             method = "lm",
  #             se = FALSE,
  #             color = "black",
  #             linetype = 1,
  #             alpha = 0.1) + 
  # scale_color_gradient(low = "#382B01",
  #                      high = "#F4E3A7") + 
  theme_light() + 
  scale_y_continuous(limits = c(-0.05, 0.4)) + 
  scale_x_discrete(expand = c(0.01, 0.5)) + 
  coord_fixed(ratio = 2) + 
  facet_grid(vars(subgenus)) + 
  ggsave("relativeabun.eps")
```

summary(lm(mean_delta_ra ~ mean_ag, data = test.df))


```{r}
unique(midwestag.vars.df$year)
```

