---
title: "Ch4 | Historical Ag & Bumble Bee Abundance/Community Changes"
subtitle: "Species occurrence and risk maps"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_notebook
---

**Name:** Jeremy Hemberger
**Email:** j.hemberger.wisc@gmail.com
**Institution:** University of Wisconsin - Madison | Department of Entomology

# LOAD PACKAGES
```{r}
library(raster)
library(rgdal)
library(sf)
library(maps)
library(tigris)
library(janitor)
library(openintro)
library(rcartocolor)
library(calecopal)
library(tidyverse)
```


# LOAD DATA
```{r}
us.counties <- map_data("county")
us.counties$subregion[us.counties$subregion == "obrien"] <- "o'brien"
us.counties$subregion[us.counties$subregion == "de kalb"] <- "dekalb"
us.counties$subregion[us.counties$subregion == "du page"] <- "dupage"
us.counties$subregion[us.counties$subregion == "st clair"] <- "saint clair"
us.counties$subregion[us.counties$subregion == "la porte"] <- "laporte"
us.counties$subregion[us.counties$subregion == "st joseph"] <- "saint joseph"
us.counties$subregion[us.counties$subregion == "st louis"] <- "saint louis"
us.counties$subregion[us.counties$subregion == "st croix"] <- "saint croix"

midwestag.model.df <- read_csv("./data/midwest_agstats.csv") %>%
  mutate(state = tolower(abbr2state(state)),
         county = tolower(county))
midwestag.other.df <- read_csv("./data/midwest_agstats_other.csv", guess_max = 5000) %>%
  dplyr::select(year, state, county, prop_pasture, prop_pest)

midwestag.all.df <- midwestag.model.df %>%
  left_join(midwestag.other.df,
            by = c("county" = "county",
                   "state" = "state", 
                   "year" = "year"))

midwestag.map.df <- midwestag.all.df %>%
  left_join(us.counties,
            by = c("state" = "region",
                   "county" = "subregion"))
```

# DRAW MAPS
## Agricultural metrics
### Map function
```{r}
plot.agstat <- function(data, 
                        agstat, 
                        nbreaks = 7, 
                        direction = 1, 
                        years = c(1880, 1900, 1950, 1974, 2017)) {
  if (direction == 1) {
    palette <- paste(carto_pal(n = nbreaks, "Fall"), sep = ",")
  } else if (direction == -1) {
    palette <- paste(rev(carto_pal(n = nbreaks, "Fall")), sep = ",")
  } else {
    print("Color palette direction must be 1 or -1")
  }
  quants <- stats::quantile(x = data[[agstat]],
                            na.rm = TRUE,
                            type = 1,
                            probs = seq(0, 1, length.out = nbreaks)) %>%
    as.vector()
  # print(quants)
  labels <- imap_chr(quants, function(., idx){
    return(paste0(round(quants[idx], 3),
                  "–",
                  round(quants[idx + 1], 3)))
  })
  labels <- labels[1:length(labels) - 1]
  # print(labels)
  data$agbrk <- cut(data[[agstat]],
                    breaks = quants,
                    include.lowest = TRUE,
                    labels = labels)
  plot <- data %>%
    filter(year %in% years) %>%
    ggplot() +
    geom_polygon(mapping = aes(x = long,
                               y = lat,
                               group = group,
                               fill = agbrk)) +
    geom_polygon(data = state_df <- map_data("state") %>%
                 filter(region %in% c("wisconsin", 
                                      "minnesota", 
                                      "michigan", 
                                      "iowa", 
                                      "illinois", 
                                      "indiana")),
               mapping = aes(x = long,
                             y = lat,
                             group = group),
               color = "#383838",
               size = 1.25,
               fill = NA) +
    scale_fill_manual(values = palette) +
    coord_map("stereographic") + 
    facet_wrap(~ year,
               ncol = 6) +
    theme_void()
  return(plot)
}

```

### Plot ag stats
```{r}
prop.crop.plot <- plot.agstat(data = midwestag.map.df,
                              agstat = "prop_cropland")
n.crop.plot <- plot.agstat(data = midwestag.map.df,
                           agstat = "n_crops",
                           nbreaks = 6,
                           direction = -1)
prop.past.plot <- plot.agstat(data = midwestag.map.df,
                              agstat = "prop_pasture",
                              direction = -1, 
                              years = c(1982, 1992, 2002, 2012))
prop.pest.plot <- plot.agstat(data = midwestag.map.df,
                              agstat = "prop_pest", 
                              years = c(1982, 1992, 2002, 2012))

prop.crop.plot
n.crop.plot
prop.past.plot
prop.pest.plot

ggsave("./model_plots/maps/prop_cropland.eps",
       prop.crop.plot,
       height = 7)
ggsave("./model_plots/maps/n_crops.eps",
       n.crop.plot,
       height = 7)
ggsave("./model_plots/maps/prop_pasture.eps",
       prop.past.plot,
       height = 7)
ggsave("./model_plots/maps/prop_pest.eps",
       prop.pest.plot, 
       height = 7)
```

## Predicted species occurrence
### Generate predictions
```{r}
pred.relabun <- function(model.list) {
  pred.list <- list()
  years.list <- list()
  years <- unique(midwestag.model.df$year)
  for (i in 1:length(model.list)) {
    for(j in years) {
      newdata.df <- midwestag.model.df %>%
        filter(year == j) %>%
        mutate(bin_ag = year)
      predict <- predict(object = model.list[[i]],
                             newdata = newdata.df,
                             type = "response",
                             na.action = na.pass,
                             allow.new.levels = TRUE)
      predict.df <- tibble(rel_abun = predict,
                           species = names(model.list[i]),
                           county = newdata.df$county,
                           state = newdata.df$state,
                           year = j)
      pred.list[[j]] <- predict.df
    }
    years.list[[i]] <- bind_rows(pred.list)
  }
  relabun.bycty.df <- bind_rows(years.list)
  return(relabun.bycty.df)
}

relabun.bycty.df <- pred.relabun(models.full.list)
```

### Map function
```{r}
library(cartography)
library(rcartocolor)
plot.predicted <- function() {
  plot.species <- c("Bombus affinis",
                    "Bombus auricomus",
                    "Bombus bimaculatus",
                    "Bombus borealis",
                    "Bombus citrinus",
                    "Bombus fervidus",
                    "Bombus griseocollis",
                    "Bombus impatiens",
                    "Bombus pensylvanicus",
                    "Bombus rufocinctus",
                    "Bombus ternarius",
                    "Bombus terricola",
                    "Bombus vagans")
  relabun.quantiles <- stats::quantile(x = relabun.bycty.df$rel_abun,
                                       na.rm = TRUE,
                                       type = 1,
                                       probs = seq(0, 1, length.out = 8)) %>%
    as.vector()
  
  prob.labels <- imap_chr(relabun.quantiles, function(., idx){
    return(paste0(round(relabun.quantiles[idx], 3),
                  "–",
                  round(relabun.quantiles[idx + 1], 3)))
  })
  prob.labels <- prob.labels[1:length(prob.labels) - 1]
  relabun.bycty.df$prob_brks <- cut(relabun.bycty.df$rel_abun,
                                    breaks = relabun.quantiles,
                                    include.lowest = TRUE,
                                    labels = prob.labels)
  prob_brks_scale <- levels(relabun.bycty.df$prob_brks)
  bumblepred.df <- relabun.bycty.df
  assign("bumblepred.df",
         bumblepred.df,
         envir = .GlobalEnv)
  for(i in plot.species) {
    plot <- relabun.bycty.df %>%
      filter(year %in% c(1900, 1950, 2017)) %>%
      filter(species == i) %>%
      mutate(bin_ag = year) %>%
      mutate(state_full = abbr2state(state)) %>%
      left_join(ungroup(us.counties),
                by = c("county" = "subregion",
                       "state" = "region")) %>%
      filter(!is.na(species)) %>%
      ggplot() +
      geom_polygon(mapping = aes(x = long,
                                 y = lat,
                                 group = group,
                                 fill = prob_brks),
                   size = 0) +
      geom_polygon(data = state_df <- map_data("state") %>%
                     filter(region %in% c("wisconsin", 
                                          "minnesota", 
                                          "michigan", 
                                          "iowa", 
                                          "illinois", 
                                          "indiana")),
                   mapping = aes(x = long,
                                 y = lat,
                                 group = group),
                   color = "#383838",
                   size = 1.25,
                   fill = NA) +
      # Plot occurrences to visualize "fit"
      # geom_point(inherit.aes = FALSE,
      #            data = spatial.bumbles.df %>%
      #              filter(species == i) %>%
      #              filter(bin_ag %in% c(1850, 1900, 1950, 1974, 2012)) %>%
      #              group_by(bin_ag) %>%
      #              sample_n(size = 50, replace = TRUE),
      #            mapping = aes(x = dec_long,
      #                          y = dec_lat),
      #            shape = 21,
      #            fill = "tomato",
    #              color = "black",
    #            size = 1) +
    # Convex hull "range"
    # geom_polygon(data = analysis.bumbles.df %>%
    #                filter(species == i) %>%
    #                slice(chull(dec_long, dec_lat)),
    #              mapping = aes(x = dec_long,
    #                            y = dec_lat),
    #              color = "tomato",
    #              fill = NA,
    #              alpha = 1,
    #              size = 0.75) +
    coord_map("stereographic") +
      labs(title = i) +
      scale_fill_manual(values = paste(rev(carto_pal(n = 8, "Fall")), sep = ","),
                        breaks = prob_brks_scale,
                        name = "Probability of occurrence",
                        drop = FALSE,
                        labels = prob_brks_scale) +
      # scale_fill_manual(values = rev(viridis::inferno(8, direction = 1)),
      #                   breaks = rev(prob_brks_scale),
      #                   name = "Probability of occurrence",
      #                   drop = FALSE,
      #                   labels = rev(prob_brks_scale)) +
      # scale_fill_viridis_c(option = "inferno",
      #                      direction = 1,
      #                      ) +
      facet_grid(~ bin_ag) + 
      theme_void()
    ggsave(paste("./model_plots/maps/prediction_maps/",
                 tolower(gsub(" ", "_", i)),
                 ".eps",
                 sep = ""),
           height = 7)
    print(plot)
  }
}
```

### Run function
```{r}
plot.predicted()
```

## Risk maps
### Assign counties to risk
```{r}
relabun.analy.list <- relabun.bycty.df %>%
  filter(!species %in% c("Bombus ashtoni",
                         "Bombus fraternus",
                         "Bombus variabilis")) %>%
  group_by(species, state, county) %>%
  group_split()

ctylm.list <- list()
for (i in 1:length(relabun.analy.list)) {
  cty.lm <- lm(rel_abun ~ scale(year),
               data = relabun.analy.list[[i]])
  ctylm.df <- tidy(cty.lm) %>%
    filter(term == "scale(year)") %>%
    mutate(species = relabun.analy.list[[i]]$species[1],
           state = relabun.analy.list[[i]]$state[1],
           county = relabun.analy.list[[i]]$county[1]) %>%
    select(species, state, county, term, estimate, std.error, p.value)
  ctylm.list[[i]] <- ctylm.df
}


ctylm.df <- bind_rows(ctylm.list)
rm(relabun.analy.list)
rm(ctylm.list)
```

##### Assign risk using quantiles
```{r}
ctylm.species.list <- ctylm.df %>%
  group_by(species) %>%
  group_split()

for (i in 1:length(ctylm.species.list)) {
  quants <- quantile(x = ctylm.df$estimate,
                     na.rm = TRUE,
                     type = 1,
                     probs = seq(0, 1, length.out = 5)) %>%
    as.vector()
  labels <- imap_chr(quants, function(., idx){
    return(paste0(round(quants[idx], 3),
                  "–",
                  round(quants[idx + 1], 3)))
  })
  labels <- labels[1:length(labels) - 1]
  ctylm.species.list[[i]]$risk_brks <- cut(ctylm.species.list[[i]]$estimate,
                            breaks = quants,
                            include.lowest = TRUE,
                            labels = ctytrend.labels)
}
ctylm.risk.df <- bind_rows(ctylm.species.list)
rm(ctylm.species.list)
```

### Map risk
```{r}
ctylm.df %>% 
  mutate(risk =  factor(as.numeric(risk_brks))) %>%
  left_join(ungroup(us.counties),
            by = c("county" = "subregion",
                   "state" = "region")) %>%
  ggplot() + 
  geom_polygon(mapping = aes(x = long,
                             y = lat,
                             group = group,
                             fill = risk)) +
  geom_polygon(data = state_df <- map_data("state") %>%
                 filter(region %in% c("wisconsin", 
                                      "minnesota", 
                                      "michigan", 
                                      "iowa", 
                                      "illinois", 
                                      "indiana")),
               mapping = aes(x = long,
                             y = lat,
                             group = group),
               color = "#383838",
               size = 1.25,
               fill = NA) +
  scale_fill_manual(values = c("#EFD53C",
                               "#B59718",
                               "#49441A",
                               "#191716")) +
  coord_map("stereographic") +
  facet_wrap(~ species) +
  theme_void() +
  ggsave("./model_plots/maps/risk_map.eps")
```

### Predicted species trend
```{r}
plot.species <- relabun.bycty.df %>%
  filter(!species %in% c("Bombus ashtoni",
                         "Bombus fraternus",
                         "Bombus variabilis")) %>%
  distinct(species) %>%
  pull(species)

for (i in plot.species) {
  plot <- relabun.bycty.df %>%
    filter(species == i) %>%
    sample_n(size = 1000) %>%
    ggplot() + 
    geom_jitter(mapping = aes(x = year,
                              y = rel_abun),
                # alpha = 0.01,
                height = 0.02,
                width = 5) +
    geom_smooth(mapping = aes(x = year,
                              y = rel_abun),
                color = "tomato",
                method = "gam") +
    theme_minimal()
  print(plot)
  ggsave(paste0("./model_plots/trends/modelavg_", i, ".eps"),
         height = 2,
         width = 3)
}
rm(plot.species)
```


## Key maps
```{r}
library(maps)

us.map <- map("state",proj="conic",par=39.83,
              fill = TRUE)

study.map <- us.map[us.map$names] %in% c("wisconsin", 
                                 "michigan", 
                                 "minnesota", 
                                 "illinois",
                                 "indiana", 
                                 "iowa")

states %>%
  ggplot() + 
  geom_sf(data = states %>%
            filter(ID %in% c("wisconsin", 
                             "michigan", 
                             "minnesota", 
                             "illinois",
                             "indiana", 
                             "iowa")),
          fill = "tomato") +
  coord_sf(crs = "+proj=aea +lat_1=25 +lat_2=50 +lon_0=-100") + 
  theme_void() +
  ggsave("./model_plots/study_area.eps")

state_df <- map_data("state")

state_df %>%
  filter(region %in% c("wisconsin", "minnesota", "michigan", "iowa", "illinois", "indiana")) %>%
  ggplot() + 
  geom_polygon(mapping = aes(x = long,
                             y = lat,
                             group = group)) + 
  coord_map("stereographic") +
  theme_void() + 
  ggsave("./model_plots/study_states.eps")
```

```{r}
devtools::install_github("an-bui/calecopal")
library(calecopal)
names(cal_palettes)

cal_palette(name = "seagrass", n = 8, type = "continuous")
```

