---
title: "Ch4 | Historical Ag & Bumble Bee Abundance/Community Changes"
subtitle: "GLM interaction visualization"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_notebook
---

**Name:** Jeremy Hemberger
**Email:** j.hemberger.wisc@gmail.com
**Institution:** University of Wisconsin - Madison | Department of Entomology

# 1870-2018 INTERACTIONS
## Generate predictions
```{r}
predict.df.list <- list()
for (i in 1:length(species.list)) {
  predict <- predictorEffects(glm.output.list[[i]],
                              se = list(compute = TRUE))
  predict.df <- bind_rows(as_tibble(predict$n_crops) %>%
                            rename(new_x = n_crops) %>%
                            mutate(var = "n_crops"), 
                          as_tibble(predict$prop_cropland) %>%
                            rename(new_x = prop_cropland) %>%
                            mutate(var = "prop_cropland"),
                          as_tibble(predict$bin_ag) %>%
                            rename(new_x = bin_ag) %>%
                            mutate(var = "bin_ag")) %>%
    mutate(species = species.list[[i]])
  predict.df.list[[i]] <- predict.df
}

all.spp.predict.df <- bind_rows(predict.df.list)
rm(predict.df.list)
```

## Plot interactions
### Cross-sectional overlay
#### Rare species
```{r}
all.spp.predict.df %>%
  filter(var == "bin_ag" &
         species %in% c("Bombus affinis", "Bombus pensylvanicus", "Bombus terricola")) %>%
  ggplot() +
  geom_ribbon(mapping = aes(x = new_x,
                            ymin = lower,
                            ymax = upper,
                            group = factor(n_crops),
                            fill = factor(n_crops)),
              alpha = 0.3) +
  geom_line(mapping = aes(x = new_x,
                          y = fit,
                          color = factor(n_crops),
                          group = factor(n_crops)),
            size = 1.25) +
  # scale_color_viridis_d(option = "inferno") +
  # scale_fill_viridis_d(option = "inferno") +
  scale_fill_manual(values = c("#080703",
                               "#6B5B0C",
                               "#D9BA1E",
                               "#F7D528",
                               "#FCE260")) +
  scale_color_manual(values = c("#080703",
                                "#6B5B0C",
                                "#D9BA1E",
                                "#F7D528",
                                "#FCE260")) +
  ylab("Predicted relative abundance") +
  xlab("Proportion cropland") +
  facet_wrap(species ~ prop_cropland,
             ncol = 5) + 
  theme_histbumbles()
  theme(strip.text.x = element_blank()) +
  ggsave("./model_plots/interactions/rare_spp.eps",
         width = 5.5,
         height = 3.5)
```

#### Common species
```{r}
all.spp.predict.df %>%
  filter(var == "bin_ag" &
         species %in% c("Bombus bimaculatus", "Bombus griseocollis", "Bombus impatiens") &
           n_crops %in% c(7, 11, 15)) %>%
  ggplot() +
  geom_ribbon(mapping = aes(x = new_x,
                            ymin = lower,
                            ymax = upper,
                            group = factor(n_crops),
                            fill = factor(n_crops)),
              alpha = 0.3) +
  geom_line(mapping = aes(x = new_x,
                          y = fit,
                          color = factor(n_crops),
                          group = factor(n_crops)),
            size = 1.25) +
  # scale_color_viridis_d(option = "inferno") +
  # scale_fill_viridis_d(option = "inferno") +
  scale_fill_manual(values = c("#080703",
                               "#6B5B0C",
                               "#D9BA1E",
                               "#F7D528",
                               "#FCE260")) +
  scale_color_manual(values = c("#080703",
                                "#6B5B0C",
                                "#D9BA1E",
                                "#F7D528",
                                "#FCE260")) +
  ylab("Predicted relative abundance") +
  xlab("Proportion cropland") +
  facet_wrap(species ~ prop_cropland,
             ncol = 5,
             ) + 
  theme_histbumbles() +
  theme(strip.text.x = element_blank()) +
  ggsave("./model_plots/interactions/common_spp.eps",
         width = 5.5,
         height = 3.5)
```

#### Other species
```{r}
all.spp.predict.df %>%
  filter(var == "bin_ag" &
         !species %in% c("Bombus bimaculatus", 
                         "Bombus griseocollis", 
                         "Bombus impatiens",
                         "Bombus affinis",
                         "Bombus pensylvanicus",
                         "Bombus terricola",
                         "Bombus ashtoni",
                         "Bombus fraternus",
                         "Bombus variabilis") &
           n_crops %in% c(7, 11, 15) &
           prop_cropland %in% c(0.2, 0.5, 0.9)) %>%
  ggplot() +
  geom_ribbon(mapping = aes(x = new_x,
                            ymin = lower,
                            ymax = upper,
                            group = factor(n_crops),
                            fill = factor(n_crops)),
              alpha = 1) +
  geom_line(mapping = aes(x = new_x,
                          y = fit,
                          color = factor(n_crops),
                          group = factor(n_crops)),
            size = 1.25) +
  scale_fill_manual(values = c("#080703",
                               "#6B5B0C",
                               "#D9BA1E",
                               "#F7D528",
                               "#FCE260")) +
  scale_color_manual(values = c("#080703",
                                "#6B5B0C",
                                "#D9BA1E",
                                "#F7D528",
                                "#FCE260")) +
  ylab("Predicted relative abundance") +
  xlab("Proportion cropland") +
  facet_wrap(species ~ prop_cropland,
             ncol = 3) + 
  theme_histbumbles() +
  ggsave("./model_plots/interactions/other_spp.eps",
         width = 7,
         height = 14)
```

#### Ag variables
```{r}
all.spp.predict.df %>%
  filter(var == "bin_ag" & 
           species %in% c("Bombus bimaculatus", "Bombus griseocollis", "Bombus impatiens")) %>%
  ggplot() +
  geom_ribbon(mapping = aes(x = new_x,
                            ymin = lower,
                            ymax = upper,
                            group = factor(n_crops),
                            fill = factor(n_crops)),
              alpha = 0.2) +  
  geom_line(mapping = aes(x = new_x,
                          y = fit,
                          group = factor(n_crops),
                          color = factor(n_crops))) +
  facet_wrap(~species + prop_cropland,
             ncol = 5) +
  scale_color_viridis_d(option = "inferno") +
  scale_fill_viridis_d(option = "inferno") +
  theme_histbumbles() +
  ggsave("./model_plots/interactions/other_spp.eps",
         width = 7,
         height = 14)
```

### AIC comparison (3-way interaction)
```{r}
AIC(glm.output.list[[1]]) # 2547.754 for model w/out 3-way interaction

df2 <- tibble()
for (i in 1:length(glm.output.list)) {
  AIC <- AIC(glm.output.list[[i]])
  # print(AIC)
  df2 <- rbind(df2, AIC) %>%
    rename(aic.2way = 1)
}
df2 <- df2 %>%
  mutate(species = unique(spatial.bumbles.glmvars.df$species))

df3 <- tibble()
for (i in 1:length(glm.output.list)) {
  AIC <- AIC(glm.output.list[[i]])
  # print(AIC)
  df3 <- rbind(df3, AIC) %>%
    rename(aic.3way = 1)
}
df3 <- df3 %>%
  mutate(species = unique(spatial.bumbles.glmvars.df$species))

df2 %>%
  left_join(df3, by = "species") %>%
  mutate(delta = aic.3way - aic.2way)
```

### R-squared
```{r}
model.rsquared <- tibble()
for (i in 1:length(glm.output.list)) {
  r2 <- broom::glance(summ(glm.output.list[[i]]))$pseudo.r.squared.mcfadden
  model.rsquared <- rbind(model.rsquared, r2) %>%
    rename(pseudo.r2 = 1)
}
model.rsquared$species <- unique(spatial.bumbles.glmvars.df$species)
```

### Surface plots
#### Common species
```{r}

```

#### Rare species
```{r}

```


# 1984-2018 INTERACTIONS
## Generate predictions
```{r}
predict.df.list <- list()
species.list <- as.list(unique(spatial.bumbles.glmvars2.df$species))
for (i in 1:length(species.list)) {
  predict <- predictorEffects(glm.output2.list[[i]],
                              se = list(compute = TRUE))
  predict.df <- bind_rows(as_tibble(predict$prop_pest) %>%
                            rename(new_x = prop_pest) %>%
                            mutate(var = "prop_pest"), 
                          as_tibble(predict$prop_pasture) %>%
                            rename(new_x = prop_pasture) %>%
                            mutate(var = "prop_pasture"),
                          as_tibble(predict$bin_ag) %>%
                            rename(new_x = bin_ag) %>%
                            mutate(var = "bin_ag")) %>%
    mutate(species = species.list[[i]])
  predict.df.list[[i]] <- predict.df
}

all.spp.predict2.df <- bind_rows(predict.df.list)
rm(predict.df.list)
rm(species.list)
```

## Plot iteractions
### Rare species
```{r}
species.list <- as.list(unique(all.spp.predict2.df$species))
for (i in 1:length(species.list)) {
  plot <- all.spp.predict2.df %>%
    filter(var == "bin_ag" &
             species == species.list[[i]] &
             n_crops %in% c(3.2, 9.8, 13) &
             prop_pest %in% c(0.0, 0.2, 0.4) &
             prop_pasture %in% c(0.00, 0.04, 0.08)) %>%
    ggplot() +
    geom_ribbon(mapping = aes(x = new_x,
                              ymin = lower,
                              ymax = upper,
                              group = factor(n_crops),
                              fill = factor(n_crops)),
                alpha = 0.3) +
    geom_line(mapping = aes(x = new_x,
                            y = fit,
                            color = factor(n_crops),
                            group = factor(n_crops)),
              size = 1.25) +
    scale_color_viridis_d(option = "inferno") +
    scale_fill_viridis_d(option = "inferno") +
    ylab("Predicted relative abundance") +
    xlab("Year") +
    ggtitle(paste(species.list[[i]])) +
    facet_wrap(~ prop_pest + prop_pasture,
               ncol = 3) + 
    theme_histbumbles()
  print(plot)
}

all.spp.predict2.df %>%
  filter(var == "bin_ag" &
         # species %in% c("Bombus affinis", "Bombus pensylvanicus", "Bombus terricola") &
           n_crops %in% c(3.2, 9.8, 13) &
           prop_pest %in% c(0.0, 0.2, 0.4) &
           prop_pasture %in% c(0.00, 0.04, 0.08)) %>%
  ggplot() +
  geom_ribbon(mapping = aes(x = new_x,
                            ymin = lower,
                            ymax = upper,
                            group = factor(n_crops),
                            fill = factor(n_crops)),
              alpha = 0.3) +
  geom_line(mapping = aes(x = new_x,
                          y = fit,
                          color = factor(n_crops),
                          group = factor(n_crops)),
            size = 1.25) +
  scale_color_viridis_d(option = "inferno") +
  scale_fill_viridis_d(option = "inferno") +
  ylab("Predicted relative abundance") +
  xlab("Year") +
  facet_wrap(species ~ prop_pest + prop_pasture,
             ncol = 9) + 
  theme_histbumbles()
  ggsave("./model_plots/interactions/rare1987_spp.eps",
         width = 7,
         height = 18)

all.spp.predict2.df %>%
  filter(var == "bin_ag" &
         species %in% c("Bombus affinis", "Bombus pensylvanicus", "Bombus terricola")) %>%
  distinct(prop_pest)

all.spp.predict2.df %>%
  filter(var == "bin_ag") %>%
  distinct(prop_pest)
```

### Common species
```{r}
all.spp.predict2.df %>%
  filter(var == "bin_ag" &
         species %in% c("Bombus bimaculatus", "Bombus griseocollis", "Bombus impatiens") &
           n_crops %in% c(3.2, 9.8, 13) &
           prop_pest %in% c(0.059, 0.120, 0.230) &
           prop_pasture %in% c(0.01, 0.03, 0.05)) %>%
  ggplot() +
  geom_ribbon(mapping = aes(x = new_x,
                            ymin = lower,
                            ymax = upper,
                            group = factor(n_crops),
                            fill = factor(n_crops)),
              alpha = 1) +
  geom_line(mapping = aes(x = new_x,
                          y = fit,
                          color = factor(n_crops),
                          group = factor(n_crops)),
            size = 1.25) +
  scale_color_viridis_d(option = "inferno") +
  scale_fill_viridis_d(option = "inferno") +
  ylab("Predicted relative abundance") +
  xlab("Year") +
  facet_wrap(species ~ prop_pest + prop_pasture,
             ncol = 3) + 
  theme_histbumbles() +
  ggsave("./model_plots/interactions/common1987_spp.eps",
         width = 7,
         height = 18)

all.spp.predict2.df %>%
  filter(var == "bin_ag" &
         species %in% c("Bombus affinis", "Bombus pensylvanicus", "Bombus terricola")) %>%
  distinct(n_crops)
```

# FODDER
## Extract yearly coefficient estimates
For each species, extract coefficient of rel_abun ~ prop_cropland and rel_abun ~ n_crops for each year.
```{r}
species.list <- as.list(unique(spatial.bumbles.glmvars.df$species))
bin.ag.list <- as.list(unique(spatial.bumbles.glmvars.df$bin_ag))
glm.list <- list()
df.list <- list()
for (i in species.list) {
  for(j in bin.ag.list) {
    model.df <- spatial.bumbles.glmvars.df %>%
      filter(species == i) %>%
      filter(bin_ag == j) %>%
      filter(!is.na(species)) %>%
      filter(!is.na(prop_cropland))
    model.df.list[[j]]
    glm.model <- glm(rel_abun ~
                       scale(n_crops) + 
                       scale(prop_cropland),
                       # avgfarm + 
                       # cloverac + 
                       # pastureac + 
                     data = model.df,
                     weights = model.df$bin_ag_abundance,
                     na.action = na.exclude,
                     family = binomial(link = "logit"))
    glm.df <- tidy(glm.model) %>%
      mutate(n_records = nrow(glm.model$model)) %>%
      dplyr::select(term, 
                    n_records, 
                    everything())
    # glmqq.plot <- qqPlot(residuals(glm.model))
    # resid.plot <- plot(glm.model)
    # glm.list[[i]] <- glm.model
    df.list[[i]] <- glm.df
  }
}

test.df <- bind_rows(df.list, .id = "species")
```

## County-level rate of change
### Exploratory plot
```{r}
midwestag.vars.df %>%
  ggplot() + 
  geom_smooth(mapping = aes(x = year,
                            y = n_crops,
                            group = county,
                            color = county),
              method = "glm",
              se = FALSE) + 
  theme_histbumbles() +
  theme(legend.position = "none")

cty.list <- as.list(unique(midwestag.vars.df$county))
cty.ncrop.df.list <- list()
cty.propc.df.list <- list()

for (i in cty.list) {
  model.df <- midwestag.vars.df %>%
    filter(county == i)
  model.lm <- lm(n_crops ~ year,
                 data = model.df)
  model.df <- tidy(model.lm)
  cty.ncrop.df.list[[i]] <- model.df
}

for (i in cty.list) {
  model.df <- midwestag.vars.df %>%
    filter(county == i)
  model.lm <- lm(prop_cropland ~ year,
                 data = model.df)
  model.df <- tidy(model.lm)
  cty.propc.df.list[[i]] <- model.df
}

cty.ncrop.df <- bind_rows(cty.ncrop.df.list, .id = "county")
cty.propc.df <- bind_rows(cty.propc.df.list, .id = "county")

# Rate of change of each county in number of crops
cty.ncrop.slopes.df <- cty.ncrop.df %>%
  filter(term == "year")

# rate of change of each county in proportion cropland
cty.propc.slopes.df <- cty.propc.df %>%
  filter(term == "year")

cty.propc.slopes.df %>%
  ggplot() + 
  geom_density(mapping = aes(x = estimate),
               fill = "tomato",
               alpha = 0.4) + 
  theme_histbumbles()

midwestag.vars.df %>%
  filter(county == "Adams") %>%
  ggplot() +
  geom_point(mapping = aes(x = year,
                           y = prop_cropland,
                           color = state)) + 
  geom_smooth(mapping = aes(x = year,
                            y = prop_cropland,
                            group = state),
              method = "lm") +
  theme_histbumbles()


cty.bumble.roc.list <- spatial.bumbles.ra.df %>%
  group_by(species, county, state) %>%
  filter(n() > 5) %>%
  group_split()

cty.bumble.raslope.list <- list()
listnumber <- 0
for (i in cty.bumble.roc.list) {
  listnumber <- listnumber + 1
  model.lm <- lm(rel_abun ~ bin_ag,
                 data = i)
  model.df <- tidy(model.lm) %>%
    mutate(species = i$species[1],
           county = i$county[1])
  cty.bumble.raslope.list[[listnumber]] <- model.df
}

cty.bumble.raslope.df <- bind_rows(cty.bumble.raslope.list) %>%
  dplyr::select(species, county, everything()) %>%
  filter(term == "bin_ag")

cty.bumble.ag.delta.df <- cty.bumble.raslope.df %>%
  left_join(cty.propc.slopes.df %>%
              dplyr::select(county,
                            propc.slope = estimate,
                            propc.err = std.error),
            by = "county") %>%
  left_join(cty.ncrop.slopes.df %>%
              dplyr::select(county, 
                            ncrop.slope = estimate,
                            ncrop.err = std.error),
            by = "county")

cty.bumble.ag.delta.df %>%
  pivot_longer(cols = c(propc.slope, ncrop.slope),
               names_to = "ag.var",
               values_to = "coef.est") %>%
  ggplot() +
  geom_vline(mapping = aes(xintercept = 0),
             color = "black",
             linetype = "dashed") +
  geom_hline(mapping = aes(yintercept = 0),
             color = "black",
             linetype = "dashed") +
  geom_point(mapping = aes(x = coef.est,
                           y = estimate,
                           color = species),
             size = 4) +
  geom_smooth(mapping = aes(x = coef.est,
                            y = estimate,
                            group = species,
                            color = species),
              method = "lm",
              se = FALSE,
              size = 2) + 
  scale_color_viridis_d(option = "inferno") +
  facet_wrap(~ag.var, scales = "free") +
  theme_histbumbles()
```

```{r}
all.spp.predict.df %>%
  filter(var == "n_crops" & 
           species %in% c("Bombus affinis", "Bombus pensylvanicus", "Bombus terricola")) %>%
  group_by(prop_cropland, bin_ag, new_x, species) %>%
  slice(1)
  ggplot() +
  # geom_ribbon(mapping = aes(x = new_x,
  #                           ymin = lower,
  #                           ymax = upper,
  #                           group = factor(n_crops),
  #                           fill = factor(n_crops)),
  #             alpha = 0.2) +  
  geom_point(mapping = aes(x = prop_cropland,
                          y = n_crops,
                          group = factor(new_x),
                          color = factor(new_x))) +
  # geom_line(mapping = aes(x = new_x,
  #                         y = n_crops,
  #                         group = factor(bin_ag),
  #                         color = factor(bin_ag))) +
  facet_wrap(~species) +
  scale_color_viridis_d(option = "inferno") +
  scale_fill_viridis_d(option = "inferno") +
  theme_histbumbles()
```



```{r}
all.spp.predict.df %>%
  filter(var == "bin_ag" &
         species %in% c("Bombus bimaculatus", "Bombus griseocollis", "Bombus impatiens")) %>%
  ggplot() +
  geom_ribbon(mapping = aes(x = new_x,
                            ymin = lower,
                            ymax = upper,
                            group = factor(n_crops),
                            fill = factor(n_crops)),
              alpha = 1) +
  geom_line(mapping = aes(x = new_x,
                          y = fit,
                          color = factor(n_crops),
                          group = factor(n_crops)),
            size = 1.25) +
  # scale_color_viridis_d(option = "inferno") +
  # scale_fill_viridis_d(option = "inferno") +
  scale_fill_manual(values = c("#080703",
                               "#6B5B0C",
                               "#D9BA1E",
                               "#F7D528",
                               "#FCE260")) +
  scale_color_manual(values = c("#080703",
                                "#6B5B0C",
                                "#D9BA1E",
                                "#F7D528",
                                "#FCE260")) +
  ylab("Predicted relative abundance") +
  xlab("Proportion cropland") +
  facet_wrap(species ~ prop_cropland,
             ncol = 5,
             ) + 
  theme_histbumbles() +
  theme(strip.text.x = element_blank())

all.spp.predict.df %>%
  filter(var == "bin_ag" &
         species %in% c("Bombus affinis", "Bombus pensylvanicus", "Bombus terricola")) %>%
  ggplot() +
  geom_ribbon(mapping = aes(x = new_x,
                            ymin = lower,
                            ymax = upper,
                            group = factor(n_crops),
                            fill = factor(n_crops)),
              alpha = 1) +
  geom_line(mapping = aes(x = new_x,
                          y = fit,
                          color = factor(n_crops),
                          group = factor(n_crops)),
            size = 1.25) +
  # scale_color_viridis_d(option = "inferno") +
  # scale_fill_viridis_d(option = "inferno") +
  scale_fill_manual(values = c("#080703",
                               "#6B5B0C",
                               "#D9BA1E",
                               "#F7D528",
                               "#FCE260")) +
  scale_color_manual(values = c("#080703",
                                "#6B5B0C",
                                "#D9BA1E",
                                "#F7D528",
                                "#FCE260")) +
  ylab("Predicted relative abundance") +
  xlab("Proportion cropland") +
  facet_wrap(species ~ prop_cropland,
             ncol = 5,
             ) + 
  theme_histbumbles() +
  theme(strip.text.x = element_blank())
```


##Try out `visreg` package
```{r}
library(visreg)

for (i in 1:length(glm.output.list)) {
  print(names(glm.output.list[i]))
  plot <- visreg2d(glm.output.list[[i]], 
           "prop_cropland", 
           "n_crops", 
           scale = "response",
           #main = names(glm.output.list[i]))
           plot.type = "gg")
  print(plot)
}

visreg2d(glm.output.list[[1]], 
         "prop_cropland", 
         "n_crops",
         # cond = list(bin_ag = 1900),
         scale = "response",
         plot.type = "gg") +
  geom_point(data = midwestag.vars.df %>%
            filter(year == 1900),
          mapping = aes(x = prop_cropland,
                        y = n_crops)) +
  ggtitle("B. impatiens: 1900")
# visreg2d(glm.output.list[[1]], 
#          "prop_cropland", 
#          "n_crops",
#          cond = list(bin_ag = 1950),
#          scale = "response")
# visreg2d(glm.output.list[[1]], 
#          "prop_cropland", 
#          "n_crops",
#          cond = list(bin_ag = 1972),
#          scale = "response")
visreg2d(glm.output.list[[1]], 
         "prop_cropland", 
         "n_crops",
         # cond = list(bin_ag = 2017),
         scale = "response",
         plot.type = "gg")  +
  geom_point(data = midwestag.vars.df %>%
            filter(year == 2017),
          mapping = aes(x = prop_cropland,
                        y = n_crops)) +
  ggtitle("B. impatiens: 2017")

```
## New model framework? glmer
```{r}
species.list <- spatial.bumbles.glmvars.df %>%
  distinct(species) %>%
  filter(!species %in% c("Bombus ashtoni",
                         "Bombus fraternus",
                         "Bombus variabilis")) %>%
  pull() %>%
  as.list()
glmer.df.list <- list()
glmer.output.list <- list()
for (i in 1:length(species.list)) {
  model.df <- spatial.bumbles.glmvars.df %>%
          filter(species == species.list[[i]])
  glmer <- glmer(rel_abun ~ scale(prop_cropland)*scale(n_crops) + (1 | bin_ag),
        data = model.df,
        weights = bin_ag_abundance,
        na.action = na.exclude,
        family = binomial(link = "logit"),
        control=glmerControl(optimizer="bobyqa",
                                 optCtrl=list(maxfun=2e5))) 
  glmer.output.list[[i]] <- glmer
  glmer.df <- tidy(glmer) %>%
      mutate(n_records = nrow(model.df),
             species = species.list[[i]]) %>%
      dplyr::select(species,
                    term, 
                    n_records, 
                    everything())
  print(glmer.df)
  glmer.df.list[[i]] <- glmer.df
}  
names(glmer.output.list) <- species.list
glmers.df <- bind_rows(glmer.df.list)

visreg2d(glmer.output.list[['Bombus impatiens']], 
         "prop_cropland", 
         "n_crops",
         # cond = list(bin_ag = 1900),
         scale = "response",
         plot.type = "gg")  +
  geom_jitter(data = midwestag.vars.df %>%
            filter(year == 1900),
          mapping = aes(x = prop_cropland,
                        y = n_crops)) +
  ggtitle("B. terricola: Ag stats 1900")


visreg2d(glmer.output.list[['Bombus impatiens']], 
         "prop_cropland", 
         "n_crops",
         # cond = list(bin_ag = 2017),
         scale = "response",
         plot.type = "gg",
         re.form = NULL,
         print.cond = TRUE) +
  geom_jitter(data = midwestag.vars.df %>%
                # filter(species == "Bombus terricola") %>%
                filter(year == 2017),
          mapping = aes(x = prop_cropland,
                        y = n_crops)) +
  ggtitle("B. terricola: Ag stats 2017")

for (i in 1:length(glmer.output.list)) {
  visreg2d(glmer.output.list[[i]], 
         "prop_cropland", 
         "n_crops",
         #cond = list(bin_ag = 2017),
         scale = "response",
         plot.type = "gg",
         allow.new.levels = TRUE)
}


plot(predictorEffects(glm.test.list[['Bombus terricola']]), "bin_ag")


```

```{r}
glm.df.list <- list()
glm.aug.list <- list()
glm.test.list <- list()
for (i in 1:length(species.list)) {
  model.df <- spatial.bumbles.glmvars.df %>%
          filter(species == species.list[[i]])
  glm <- glm(rel_abun ~ prop_cropland*n_crops*bin_ag,
        data = model.df,
        weights = model.df$bin_ag_abundance,
        na.action = na.exclude,
        family = binomial(link = "logit")) 
  glm.test.list[[i]] <- glm
  glm.model.df <- augment(glm, 
                          newdata = model.df)
  glm.df <- tidy(glm) %>%
      mutate(n_records = nrow(model.df),
             species = species.list[[i]]) %>%
      dplyr::select(species,
                    term, 
                    n_records, 
                    everything())
  # print(glmer.df)
  glm.df.list[[i]] <- glm.df
  glm.aug.list[[i]] <- glm.model.df
}  
names(glm.test.list) <- species.list
glms.test.df <- bind_rows(glm.df.list)
glms.aug.df <- bind_rows(glm.aug.list)

library(jtools)
library(interactions)
  

species.df.list <- spatial.bumbles.glmvars.df %>%
  filter(!is.na(rel_abun) &
           !is.na(prop_cropland) &
           !is.na(n_crops)) %>%
  group_by(species) %>%
  group_split()
names(species.df.list) <- unique(spatial.bumbles.glmvars.df$species)

n_crops*prop_cropland*bin_ag
prop_cropland*n_crops*bin_ag

glm <- glm(rel_abun ~ n_crops*prop_cropland*bin_ag,
           data = species.df.list[['Bombus terricola']],
           weights = bin_ag_abundance,
           family = binomial(link = "logit"))


           
interact_plot(sma.models.list[[1]], 
              pred = bin_ag, 
              modx = n_crops, 
              mod2 = prop_cropland, 
              # mod2.values = c(0.2, 0.5, 0.8),
              # modx.values = c(3, 7, 15),
              interval = TRUE) +
  theme_histbumbles()

probe_interaction(glm, 
              pred = bin_ag, 
              modx = n_crops, 
              mod2 = prop_cropland, 
              interval = TRUE)
plot(sim_slopes(glm, 
              pred = bin_ag, 
              modx = n_crops, 
              mod2 = prop_cropland, 
              interval = TRUE))
library(rgl)
visreg2d(glm, "prop_cropland", "n_crops", plot.type = "persp", scale = "response")
visreg2d(glm, "prop_cropland", "n_crops", scale = "response")
```

