---
title: "Ch4 | Historical Ag & Bumble Bee Abundance/Community Changes"
subtitle: "GLM model fit and visualization"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_notebook
---

**Name:** Jeremy Hemberger

**Email:** [j.hemberger.wisc\@gmail.com](mailto:j.hemberger.wisc@gmail.com){.email}

**Institution:** University of Wisconsin - Madison \| Department of Entomology

# LOAD PACKAGES

```{r}
library(jtools)
library(nlme)
library(lme4)
library(broom.mixed)
library(interactions)
library(janitor)
library(huxtable)
library(ggpubr)
library(tidyverse)
```

# LOAD DATA

```{r}
model.vars.df <- read_csv("./data/model_1_bumbles.csv") %>%
  rename(bumble_abun = bumble.abun)
model.vars.uniq.df <- read_csv("./data/model_2_bumbles.csv") %>%
  rename(bumble_abun = bumble.abun)
model.vars.3yr.df <- read_csv("./data/model_3_bumbles.csv") %>%
  rename(bumble_abun = bumble.abun)
```

# PREP DATA

## Species data frames

```{r}
species.list <- as.list(unique(model.vars.df$species))
species.df.list <- model.vars.df %>%
  group_by(species) %>%
  group_split()
names(species.df.list) <- species.list
```

# MODEL: 1870-2017

## Function: fit models

```{r}
fit.glms <- function(data) {
  glm.df.list <- list()
  glm.list <- list()
  for (i in 1:length(species.list)) {
    model.df <- data %>%
      filter(species == species.list[[i]])
    glm <- glm(rel_abun ~ prop_cropland*n_crops*bin_ag,
               data = model.df,
               weights = bumble_abun,
               na.action = na.exclude,
               # family = binomial(link = "logit"),
               family = quasibinomial(link = "logit")
               ) 
    glm.list[[i]] <- glm
    # glm.df <- tidy(glm) %>%
    #   mutate(n_records = nrow(model.df),
    #          species = unique(model.df$species),
    #          r2 = broom::glance(summ(glm))$pseudo.r.squared.mcfadden) %>%
    #   dplyr::select(species,
    #                 term, 
    #                 n_records,
    #                 r2,
    #                 everything())
    # glm.df.list[[i]] <- glm.df
  }
  names(glm.list) <- species.list
  # glms.df <- bind_rows(glm.df.list)

  # end.list <- list(full.model.list = glm.list, full.model.df = glms.df)
  return(glm.list)
}
```

## Function: scale and extract

```{r}
scale_coefs <- function(model.list) {
  df.list <- list()
  for (i in 1:length(model.list)) {
    species <- names(model.list[i])
    summ.glm <- summ(model.list[[i]],
                     scale = TRUE,
                     confint = TRUE)
    # print(summ.glm)
    glm.df <- as.data.frame(summ.glm[["coeftable"]]) %>%
      mutate(term = c("(Intercept)",
                      "prop_cropland",
                      "n_crops",
                      "bin_ag",
                      "prop_cropland:n_crops",
                      "prop_cropland:bin",
                      "n_crops:bin_ag",
                      "prop_cropland:n_crops:bin_ag")) %>%
      dplyr::select("term" = "term",
                    "estimate" = "Est.",
                    "lower.ci" = "2.5%",
                    "upper.ci" = "97.5%",
                    "z.stat" = "z val.",
                    "p.value" = "p") %>%
      mutate(species = species,
             sample.size = length(summ.glm$model$y),
             r2 = broom::glance(summ(model.list[[i]]))$pseudo.r.squared.mcfadden,
             p.value = ifelse(p.value < 0.001,
                              "< 0.001",
                              round(p.value,
                                    digits = 3)),
             estimate = round(estimate,
                              digits = 3),
             lower.ci = round(lower.ci,
                              digits = 4),
             upper.ci = round(upper.ci,
                              digits = 4)) %>%
      dplyr::select(species, sample.size, everything())
    # print(glm.df)
    df.list[[i]] <- glm.df
  }
  glms.df <- bind_rows(df.list)
  scaled.model.list <- list(scaled.glms.df = glms.df)
  return(scaled.model.list)
}
```

## Function: visualize interactions

```{r}
plot.interx <- function(model.list) {
  plot.list <- list()
  slope.test.list <- list()
  slope.list <- list()
  table.list <- list()
  for (i in 1:length(model.list)) {
    plot <- interactions::interact_plot(model.list[[i]],
                                        pred = prop_cropland,
                                        modx = n_crops,
                                        mod2 = bin_ag,
                                        outcome.scale = "response",
                                        interval = TRUE,
                                        # x.label = "Year",
                                        x.label = "Proportion agriculture",
                                        y.label = "Probability of occurrence",
                                        mod2.values = c(1870, 1950, 2017),
                                        modx.values = c(5, 10, 15),
                                        modx.labels = c("5", "10", "15"), 
                                        # modx.labels = c("Mean crops - 1 S.D.",
                                        #                 "Mean crops",
                                        #                 "Mean crops + 1 S.D."),
                                        mod2.labels = c("1870",
                                                        "1900",
                                                        "2012"))
    plot.df <- ggplot_build(plot)$plot$data
    
    plot <- plot.df %>%
      # mutate(modx_group = factor(modx_group,
      #                            levels = c("Mean crops",
      #                                       "Mean crops - 1 S.D.",
      #                                       "Mean crops + 1 S.D."))) %>%
      mutate(modx_group = factor(modx_group,
                                 levels = c(5, 10, 15))) %>%
      ggplot() + 
      geom_ribbon(mapping = aes(x = prop_cropland,
                                y = rel_abun,
                                ymax = ymax,
                                ymin = ymin,
                                fill = modx_group),
                  alpha = 0.2) +
      geom_line(mapping = aes(x = prop_cropland,
                              y = rel_abun,
                              color = modx_group,
                              linetype = modx_group),
                size = 1.5) +
      scale_color_manual(values = c("#000000",
                                    "#CC3333",
                                    "#669966")) +
      scale_fill_manual(values = c("#000000",
                                    "#CC3333",
                                    "#669966")) +
      # scale_y_continuous(limits = c(0, 1),
      #                    breaks = seq(0, 1, by = 0.25)) +
      scale_x_continuous(limits = c(0, 1),
                         breaks = seq(0, 1, by = 0.25)) +
      facet_wrap(~ bin_ag, scales = "free") +
      theme_nice()
      slope.test <- interactions::sim_slopes(model.list[[i]],
                                             pred = prop_cropland,
                                             modx = n_crops,
                                             mod2 = bin_ag,
                                             interval = TRUE,
                                             mod2.values = c(1870, 1950, 2017),
                                             modx.labels = c("Mean crops - 1 S.D.",
                                                             "Mean crops",
                                                             "Mean crops + 1 S.D."),
                                             mod2.labels = c("1870",
                                                             "1950",
                                                             "2012"))
                                             # mod2.labels = c("Mean cropland - 1 S.D.",
                                             #                 "Mean cropland",
                                             #                 "Mean cropland + 1 S.D."))
    slope.plot <- plot(slope.test)
    slope.table <- as_huxtable(slope.test)
    
    plot.list[[i]] <- plot
    slope.test.list[[i]] <- slope.test
    slope.list[[i]] <- slope.plot
    table.list[[i]] <- slope.table
    print(plot)
  }
  names(plot.list) <- species.list
  names(slope.test.list) <- species.list
  names(slope.list) <- species.list
  names(table.list) <- species.list
  interactions.full.list <- list(plots = plot.list,
                                 slope.tests = slope.test.list,
                                 slopes = slope.list,
                                 tables = table.list)
  return(interactions.full.list)
}
```

## Run Models

#### Fit

```{r}
models.full.list <- fit.glms(data = model.vars.df)
models.uniq.list <- fit.glms(data = model.vars.uniq.df)
models.3yr.list <- fit.glms(data = model.vars.df)
```

#### Scale and extract

```{r}
scaled.models.full.list <- scale_coefs(models.full.list)
scaled.model.uniq.list <- scale_coefs(models.uniq.list)
scaled.model.3yr.list <- scale_coefs(models.3yr.list)
```

#### Visualize interactions

```{r}
interactions.full.list <- plot.interx(models.full.list)
interactions.uniq.list <- plot.interx(models.uniq.list)
interactions.3yr.list <- plot.interx(models.3yr.list)
```


# SPATIAL MODELS
## Create spdf
```{r}
us.spdf <- raster::getData('GADM', 
                           country = 'USA', 
                           level = 2)
midwest.spdf <- subset(us.spdf, 
                       NAME_1 == "Iowa" | NAME_1 == "Illinois" | NAME_1 == "Indiana" |
                         NAME_1 == "Michigan" | NAME_1 == "Minnesota" | NAME_1 == "Wisconsin")
countycode.df <- tibble(GID_2 = midwest.spdf@data[["GID_2"]],
                        state = midwest.spdf@data[["NAME_1"]],
                        county_name = midwest.spdf@data[["NAME_2"]])

spatial.corr.test.df <- model.vars.df %>%
  left_join(countycode.df,
            by = c("state" = "state",
                   "county" = "county_name"))
spatial.corr.test.spdf <- sp::merge(midwest.spdf,
                                    spatial.corr.test.df,
                                    by = "GID_2",
                                    duplicateGeoms = TRUE)

table(spatial.corr.test.spdf@data$species)
```

## Test for spatial autocorrelation 
```{r}
spatial.autocorr.test <- function(spdf) {
  analysis.spdf <- spdf %>%
    filter(!is.na(species)) %>%
    filter(!is.na(prop_cropland))
  spp.test.yr.df <- spatial.corr.test.df %>%
    group_by(species, bin_ag) %>%
    summarise(n_counties = n()) %>%
    filter(n_counties == max(n_counties),
           !species %in% c("Bombus ashtoni",
                           "Bombus fraternus",
                           "Bombus variabilis"))
  species <- c("Bombus affinis",
               "Bombus auricomus",
               "Bombus bimaculatus",
               "Bombus borealis",
               "Bombus citrinus",
               "Bombus fervidus",
               "Bombus griseocollis",
               "Bombus impatiens",
               "Bombus pensylvanicus",
               "Bombus rufocinctus",
               "Bombus ternarius",
               "Bombus terricola",
               "Bombus vagans")
  iter <- 0
  for (i in species) {
    iter <- iter + 1
    bin_ag_max <- spp.test.yr.df$bin_ag[spp.test.yr.df$species == i]
    model.spdf <- analysis.spdf %>%
      filter(species == i) %>%
      filter(bin_ag == bin_ag_max)
    spdf.list[[iter]] <- model.spdf
  }
  names(spdf.list) <- species
  assign("spatial.autocorr.spdf.list",
         spdf.list,
         envir = .GlobalEnv)
  
  listnumber = 0
  modelresults.list <- list()
  modeloutput.list <- list()
  for (i in spdf.list) {
    listnumber <- listnumber + 1
    # OLS Model
    glm.model <- glm(rel_abun ~ scale(n_crops) * scale(prop_cropland) * scale(bin_ag),
                     data = i@data,
                     weights = i@data$bumble_abun,
                     na.action = na.exclude)
    glm.df <- tidy(glm.model) %>%
      mutate(model.type = "GLM",
             species = i@data$species[1],
             time = i@data$bin_ag[1],
             sample.size = nrow(i@data),
             moran.pvalue = NA) %>%
      dplyr::select(species, time, sample.size, model.type, moran.pvalue, everything())
    modeloutput.list[[listnumber]] <- glm.model
    # print(paste("n residuals:", length(resid(glm.model)), sep = " "))
    # Create spatial weights matrix
    neighbors <- poly2nb(i,
                         queen = TRUE)
    # print(paste("n neighbors", length(neighbors), sep = " "))
    weights <- nb2listw(neighbors,
                        # style = "W",
                        zero.policy = TRUE)
    moran.test <- lm.morantest(glm.model,
                               weights,
                               alternative = "two.sided",
                               zero.policy = TRUE)
    moran.pvalue <- moran.test$p.value[1, 1]
    glm.df$moran.pvalue = moran.pvalue
    modelresults.list[[listnumber]] <- glm.df
  }
  spatial.corr.df <- bind_rows(modelresults.list)
}

autocorr.spp.df <- spatial.autocorr.test(spatial.corr.test.spdf)
```

## Run spatial models
```{r}
spatial.corr.models <- function(spdf) {
  analysis.spdf <- spdf %>%
    filter(!is.na(species))
  spatial.test.list <- list()
  spdf.list <- list()
  species <- c("Bombus affinis",
               "Bombus impatiens",
               "Bombus ternarius",
               "Bombus vagans")
  for (i in species) {
    model.spdf <- analysis.spdf %>%
      filter(species == i) %>%
      filter(!is.na(prop_cropland) &
               !is.na(n_crops) &
               !is.na(rel_abun))
    spdf.list[[i]] <- model.spdf
    # assign("test.list",
    #        spdf.list,
    #        envir = .GlobalEnv)
  }
  listnumber <- 0
  modelresults.list <- list()
  modeloutput.list <- list()
  sma.results.list <- list()
  sma.models.list <- list()
  predict.list <- list()
  predict.glm.list <- list()
  for (i in spdf.list) {
    listnumber <- listnumber + 1
    glm.model <- glm(rel_abun ~
                       scale(n_crops) * scale(prop_cropland) * scale(bin_ag),
                     data = i@data,
                     weights = i@data$bumble_abun,
                     family = binomial(link = "logit"),
                     na.action = na.exclude)
    glm.df <- tidy(glm.model) %>%
      mutate(model.type = "glm",
             species = i@data$species[1],
             obsv = nrow(i@data)) %>%
      dplyr::select(species, model.type, obsv, everything())
    # residuals(ols.model)
    modeloutput.list[[listnumber]] <- glm.model
    weights.queen.list <- poly2nb(i,
                                  queen = TRUE)
    # print(length(weights.queen.list))
    # print(length(ols.model$residuals))
    # acf(residuals(ols.model))
    weights <- nb2listw(weights.queen.list,
                        style = "W",
                        zero.policy = TRUE)
    # plot(weights, coordinates(i))
    moran.test <- lm.morantest(glm.model,
                               weights,
                               alternative = "two.sided",
                               zero.policy = TRUE)
    moran.pvalue <- moran.test$p.value[1, 1]
    # print(moran.test)
    glm.df <- glm.df %>%
      mutate(moran_pvalue = moran.test$p.value)
    modelresults.list[[listnumber]] <- glm.df
    fit.sma <- errorsarlm(rel_abun ~
                            scale(n_crops) * scale(prop_cropland) * scale(bin_ag),
                          data = i,
                          weights = bumble_abun,
                          listw = weights,
                          zero.policy = TRUE,
                          na.action = na.omit)
    print(summary(fit.sma))
    sma.models.list[[listnumber]] <- fit.sma
    sma.model.df <- tibble(term = c("(Intercept)",
                                    "n_crops",
                                    "prop_cropland",
                                    "bin_ag",
                                    "n_crops:bin_ag",
                                    "prop_cropland:bin_ag",
                                    "n_crops:prop_cropland",
                                    "n_crops:prop_cropland:bin_ag"),
                           estimate = (summary(fit.sma)$Coef[, 1]),
                           std.error = (summary(fit.sma)$Coef[, 2]),
                           statistic = (summary(fit.sma)$Coef[, 3]),
                           p.value = (summary(fit.sma)$Coef[, 4])) %>%
      mutate(model.type = "spatial.error",
             species = i@data$species[1],
             sample.size = nrow(i@data)) %>%
      mutate(moran_pvalue = moran.test$p.value) %>%
      left_join(as_tibble(confint(fit.sma)) %>%
                  mutate(term = c("lamda",
                                  "(Intercept)",
                                  "n_crops",
                                  "prop_cropland",
                                  "bin_ag",
                                  "n_crops:bin_ag",
                                  "prop_cropland:bin_ag",
                                  "n_crops:prop_cropland",
                                  "n_crops:prop_cropland:bin_ag"
                  )) %>%
                  dplyr::select(term, lower.ci = `2.5 %`, upper.ci = `97.5 %`),
                by = c("term" = "term")) %>%
      dplyr::select(species,
                    model.type,
                    sample.size,
                    moran_pvalue,
                    term,
                    estimate,
                    lower.ci,
                    upper.ci,
                    z.stat = statistic,
                    p.value)
    print(sma.model.df)
    sma.results.list[[listnumber]] <- sma.model.df
    predict.df <- predict(fit.sma,
                          newdata = i,
                          listw = weights,
                          all.data = TRUE)
    predict.df$species <- i@data$species[1]
    predict.glm.df <- predict(glm.model,
                              newdata = i,
                              type = "response")
    predict.glm.list[[listnumber]] <- predict.glm.df
    predict.list[[listnumber]] <- predict.df
  }
  # sma.predict.df <- bind_rows(predict.list)
  glm.results.df <- bind_rows(modelresults.list)
  sma.results.df <- bind_rows(sma.results.list)
  assign("sma.predict.list",
         predict.list,
         envir = .GlobalEnv)
  assign("glm.predict.list",
         predict.glm.list,
         envir = .GlobalEnv)
  assign("glm.spatialauto.df",
         glm.results.df,
         envir = .GlobalEnv)
  assign("sma.spatialauto.df",
         sma.results.df,
         envir = .GlobalEnv)
  assign("sma.models.list",
         sma.models.list,
         envir = .GlobalEnv)
}

spatial.corr.models(spatial.corr.test.spdf)
```

## Prediction differences
```{r}
spatial.prediction.diff <- function() {
  species <- c("Bombus affinis",
               "Bombus impatiens",
               "Bombus ternarius",
               "Bombus vagans")
  df.list <- list()
  for (i in 1:length(sma.predict.list)) {
    df <- tibble(spatial.model = unlist(sma.predict.list[[i]])) %>%
      filter(!str_detect(spatial.model, 'Bombus')) %>%
      mutate(spatial.model = as.numeric(spatial.model)) %>%
      bind_cols(tibble(glm.model = unlist(glm.predict.list[[i]]))) %>%
      mutate(est.diff = spatial.model - glm.model,
             species = species[i])
    df.list[[i]] <- df
  }
  final.df <- bind_rows(df.list)
}

model.dif.df <- spatial.prediction.diff()
hist(model.dif.df$est.diff)
```

# MODEL: 1982-2012

## Function: fit models

```{r}
fit.1982.glms <- function(data) {
  species.list <- list("Bombus affinis", 
                       "Bombus bimaculatus",
                       "Bombus griseocollis",
                       "Bombus impatiens",
                       "Bombus pensylvanicus",
                       "Bombus terricola")
  glm.df.list <- list()
  glm.list <- list()
  for (i in 1:length(species.list)) {
    model.df <- data %>%
      filter(species == species.list[[i]])
    glm <- glm(rel_abun ~ scale(n_crops) * 
                       scale(prop_pasture) * 
                       scale(prop_pest) * 
                       scale(bin_ag),
               data = model.df,
               weights = bumble_abun,
               na.action = na.exclude,
               family = binomial(link = "logit")) 
    glm.list[[i]] <- glm
    # glm.df <- tidy(glm) %>%
    #   mutate(n_records = nrow(model.df),
    #          species = unique(model.df$species),
    #          r2 = broom::glance(summ(glm))$pseudo.r.squared.mcfadden) %>%
    #   dplyr::select(species,
    #                 term, 
    #                 n_records,
    #                 r2,
    #                 everything())
    # glm.df.list[[i]] <- glm.df
  }
  names(glm.list) <- species.list
  # glms.df <- bind_rows(glm.df.list)

  # end.list <- list(full.model.list = glm.list, full.model.df = glms.df)
  return(glm.list)
}
```

## Function: scale and extract

```{r}
scale.1982.coefs <- function(model.list) {
  df.list <- list()
  for (i in 1:length(model.list)) {
    species <- names(model.list[i])
    summ.glm <- summ(model.list[[i]],
                     scale = TRUE,
                     confint = TRUE)
    # print(summ.glm)
    glm.df <- as.data.frame(summ.glm[["coeftable"]]) %>%
      mutate(term = c("(Intercept)",
                      "n_crops",
                      "prop_pasture",
                      "prop_pest",
                      "bin_ag",
                      "prop_pasture:bin_ag",
                      "prop_pest:bin_ag",
                      "n_crops:bin_ag",
                      "n_crops:prop_pest",
                      "n_crops:prop_pasture",
                      "prop_pasture:prop_pest")) %>%
      dplyr::select("term" = "term",
                    "estimate" = "Est.",
                    "lower.ci" = "2.5%",
                    "upper.ci" = "97.5%",
                    "z.stat" = "z val.",
                    "p.value" = "p") %>%
      mutate(species = species,
             sample.size = length(summ.glm$model$y),
             r2 = broom::glance(summ(model.list[[i]]))$pseudo.r.squared.mcfadden,
             p.value = ifelse(p.value < 0.001,
                              "< 0.001",
                              round(p.value,
                                    digits = 3)),
             estimate = round(estimate,
                              digits = 3),
             lower.ci = round(lower.ci,
                              digits = 4),
             upper.ci = round(upper.ci,
                              digits = 4)) %>%
      dplyr::select(species, sample.size, everything())
    # print(glm.df)
    df.list[[i]] <- glm.df
  }
  glms.df <- bind_rows(df.list)
  scaled.model.list <- list(scaled.glms.df = glms.df)
  return(scaled.model.list)
}

models.1982.full.list[[1]]
```

## Function: visualize interactions

```{r}
plot.1982.interx <- function(model.list) {
  plot.list <- list()
  slope.test.list <- list()
  slope.list <- list()
  table.list <- list()
  for (i in 1:length(model.list)) {
    plot <- interactions::interact_plot(model.list[[i]],
                                        pred = bin_ag,
                                        modx = n_crops,
                                        mod2 = prop_pasture,
                                        outcome.scale = "response",
                                        interval = TRUE,
                                        x.label = "Year",
                                        y.label = "Probability of occurrence") +
      
      scale_y_continuous(limits = c(0, 1),
                         breaks = seq(0, 1, by = 0.25)) +
      scale_x_continuous(limits = c(1982, 2020))
      theme_nice() + 
      theme(legend.position = "none")
      slope.test <- interactions::sim_slopes(model.list[[i]],
                                             pred = bin_ag,
                                             modx = n_crops,
                                             mod2 = prop_pasture,
                                             interval = TRUE)
    slope.plot <- plot(slope.test)
    slope.table <- as_huxtable(slope.test)
    
    plot.list[[i]] <- plot
    slope.test.list[[i]] <- slope.test
    slope.list[[i]] <- slope.plot
    table.list[[i]] <- slope.table
    print(plot)
  }
  names(plot.list) <- species.list
  names(slope.test.list) <- species.list
  names(slope.list) <- species.list
  names(table.list) <- species.list
  interactions.full.list <- list(plots = plot.list,
                                 slope.tests = slope.test.list,
                                 slopes = slope.list,
                                 tables = table.list)
  return(interactions.full.list)
}
```

## Run Models

#### Fit

```{r}
models.1982.full.list <- fit.1982.glms(data = model.vars.df)
models.1982.uniq.list <- fit.1982.glms(data = model.vars.uniq.df)
models.1982.3yr.list <- fit.1982.glms(data = model.vars.df)
```

#### Scale and extract

```{r}
scaled.1982.models.full.list <- scale.1982.coefs(models.1982.full.list)
scaled.1982.model.uniq.list <- scale.1982.coefs(models.1982.uniq.list)
scaled.1982.model.3yr.list <- scale.1982.coefs(models.1982.3yr.list)
```

#### Visualize interactions

```{r}
interactions.1982.full.list <- plot.1982.interx(models.1982.full.list)
interactions.1982.uniq.list <- plot.interx(models.1982.uniq.list)
interactions.1982.3yr.list <- plot.interx(models.1982.3yr.list)
```


# EXPORT INTERACTION FIGURES
## Plot stack function

```{r}
export.plots <- function(plots, species, export.name) {
  print(export.name)
  export.plot <- ggarrange(plots$species[1],
                           plots$species[2],
                           plots$species[3],
                           nrow = length(species),
                           common.legend = TRUE)
  ggsave(export.name,
         export.plot)
}
```

## Species groups

### Interaction plots

```{r}
full.model.intx.spp <- list(
  common.spp = interactions.full.list$plots[c("Bombus bimaculatus",
                                               "Bombus griseocollis",
                                               "Bombus impatiens")],
  consv.spp = interactions.full.list$plots[c("Bombus affinis",
                                              "Bombus pensylvanicus",
                                              "Bombus terricola")],
  other.spp = interactions.full.list$plots[c("Bombus auricomus",
                                              "Bombus borealis",
                                              "Bombus citrinus",
                                              "Bombus fervidus",
                                              "Bombus rufocinctus",
                                              "Bombus ternarius",
                                              "Bombus vagans")],
  common.uniq.spp = interactions.uniq.list$plots[c("Bombus bimaculatus",
                                                    "Bombus griseocollis",
                                                    "Bombus impatiens")],
  consv.uniq.spp = interactions.uniq.list$plots[c("Bombus affinis",
                                                   "Bombus pensylvanicus",
                                                   "Bombus terricola")],
  other.uniq.spp = interactions.uniq.list$plots[c("Bombus auricomus",
                                                   "Bombus borealis",
                                                   "Bombus citrinus",
                                                   "Bombus fervidus",
                                                   "Bombus rufocinctus",
                                                   "Bombus ternarius",
                                                   "Bombus vagans")],
  common.3yr.spp = interactions.3yr.list$plots[c("Bombus bimaculatus",
                                                  "Bombus griseocollis",
                                                  "Bombus impatiens")],
  consv.3yr.spp = interactions.3yr.list$plots[c("Bombus affinis",
                                                 "Bombus pensylvanicus",
                                                 "Bombus terricola")],
  other.3yr.spp = interactions.3yr.list$plots[c("Bombus auricomus",
                                                 "Bombus borealis",
                                                 "Bombus citrinus",
                                                 "Bombus fervidus",
                                                 "Bombus rufocinctus",
                                                 "Bombus ternarius",
                                                 "Bombus vagans")])
```

### Interaction sig. plots

```{r}
full.model.slope.spp <- list(
  common.spp = interactions.full.list$slopes[c("Bombus bimaculatus",
                                               "Bombus griseocollis",
                                               "Bombus impatiens")],
  consv.spp = interactions.full.list$slopes[c("Bombus affinis",
                                              "Bombus pensylvanicus",
                                              "Bombus terricola")],
  other.spp = interactions.full.list$slopes[c("Bombus auricomus",
                                              "Bombus borealis",
                                              "Bombus citrinus",
                                              "Bombus fervidus",
                                              "Bombus rufocinctus",
                                              "Bombus ternarius",
                                              "Bombus vagans")],
  common.uniq.spp = interactions.uniq.list$slopes[c("Bombus bimaculatus",
                                                    "Bombus griseocollis",
                                                    "Bombus impatiens")],
  consv.uniq.spp = interactions.uniq.list$slopes[c("Bombus affinis",
                                                   "Bombus pensylvanicus",
                                                   "Bombus terricola")],
  other.uniq.spp = interactions.uniq.list$slopes[c("Bombus auricomus",
                                                   "Bombus borealis",
                                                   "Bombus citrinus",
                                                   "Bombus fervidus",
                                                   "Bombus rufocinctus",
                                                   "Bombus ternarius",
                                                   "Bombus vagans")],
  common.3yr.spp = interactions.3yr.list$slopes[c("Bombus bimaculatus",
                                                  "Bombus griseocollis",
                                                  "Bombus impatiens")],
  consv.3yr.spp = interactions.3yr.list$slopes[c("Bombus affinis",
                                                 "Bombus pensylvanicus",
                                                 "Bombus terricola")],
  other.3yr.spp = interactions.3yr.list$slopes[c("Bombus auricomus",
                                                 "Bombus borealis",
                                                 "Bombus citrinus",
                                                 "Bombus fervidus",
                                                 "Bombus rufocinctus",
                                                 "Bombus ternarius",
                                                 "Bombus vagans")])
```

## Full

### Common species

```{r}
ggarrange(plotlist = full.model.intx.spp$common.spp,
          nrow = length(full.model.intx.spp$common.spp),
          common.legend = TRUE) +
  ggsave("./model_plots/interactions/1870_common_spp.pdf",
         width = 6,
         height = 6)
```

### Consv. species

```{r}
ggarrange(plotlist = full.model.intx.spp$consv.spp,
          nrow = length(full.model.intx.spp$consv.spp),
          common.legend = TRUE) +
  ggsave("./model_plots/interactions/1870_consv_spp.pdf",
         width = 6,
         height = 6)
```

### Other species

```{r}
ggarrange(plotlist = full.model.intx.spp$other.spp,
          nrow = length(full.model.intx.spp$other.spp),
          common.legend = TRUE) +
  facet_wrap(~ ., scales = "free") +
  ggsave("./model_plots/interactions/1870_other_spp.pdf",
         width = 6,
         height = 14)
```

## Filtered

### Common species

```{r}
ggarrange(plotlist = full.model.intx.spp$common.uniq.spp,
          nrow = length(full.model.intx.spp$common.uniq.spp),
          common.legend = TRUE) +
  ggsave("./model_plots/interactions/1870_common_uniq_spp.pdf",
         width = 6,
         height = 6)
```

### Consv. species

```{r}
ggarrange(plotlist = full.model.intx.spp$consv.uniq.spp,
          nrow = length(full.model.intx.spp$consv.uniq.spp),
          common.legend = TRUE) +
  ggsave("./model_plots/interactions/1870_consv_uniq_spp.pdf",
         width = 6,
         height = 6)
```

### Other species

```{r}
ggarrange(plotlist = full.model.intx.spp$other.uniq.spp,
          nrow = length(full.model.intx.spp$other.uniq.spp),
          common.legend = TRUE) +
  ggsave("./model_plots/interactions/1870_other_uniq_spp.pdf",
         width = 6,
         height = 14)
```

## `bin_ag` windows

### Common species

```{r}
ggarrange(plotlist = full.model.intx.spp$common.3yr.spp,
          nrow = length(full.model.intx.spp$common.3yr.spp),
          common.legend = TRUE) +
  ggsave("./model_plots/interactions/1870_common_3yr_spp.pdf",
         width = 6,
         height = 6)
```

### Consv. species

```{r}
ggarrange(plotlist = full.model.intx.spp$consv.3yr.spp,
          nrow = length(full.model.intx.spp$consv.3yr.spp),
          common.legend = TRUE) +
  ggsave("./model_plots/interactions/1870_consv_3yr_spp.pdf",
         width = 6,
         height = 6)
```

### Other species

```{r}
ggarrange(plotlist = full.model.intx.spp$other.3yr.spp,
          nrow = length(full.model.intx.spp$other.3yr.spp),
          common.legend = TRUE) +
  ggsave("./model_plots/interactions/1870_other_3yr_spp.pdf",
         width = 6,
         height = 14)
```

# EXPORT SLOPE INFO

## Common species

```{r}
ggarrange(plotlist = full.model.slope.spp$common.spp,
          nrow = 1) +
  ggsave("./model_plots/interactions/1870_common_spp_slopes.eps",
         width = 8,
         height = 5)
```

## Consv. species

```{r}
ggarrange(plotlist = full.model.slope.spp$consv.spp,
          nrow = 1) +
  ggsave("./model_plots/interactions/1870_consv_spp_slopes.eps",
         width = 8,
         height = 5)
```

## Other species

```{r}
ggarrange(plotlist = full.model.slope.spp$other.spp,
          nrow = 1,
          ncol = 7) +
  ggsave("./model_plots/interactions/1870_other_spp_slopes.eps",
         width = 16,
         height = 5)

# Save slope tables
for (i in 1:length(interactions.full.list$slope.tests)) {
  quick_pdf(as_huxtable(interactions.full.list$slope.tests[[i]]) %>%
              theme_article(),
            file = paste0("./model_plots/interactions/slope_tables/",
                          make_clean_names(names(interactions.full.list$slope.tests[i])),
                          ".pdf"))
}
```

# EXPORT MODEL TABLES

```{r}
scaled.models.full.list$scaled.glms.df %>%
  write_csv("./model_output/1870_full_model.csv")
```

# TEST NO XERC DATA
```{r}
model.vars.noxerx.df <- read_csv("./data/gbif_bumbles.csv") %>%
  
```

# GLOBAL MODEL
```{r}
model.vars.df %>%
  group_by(species) %>%
  count()

global.glm <- glmer(rel_abun ~ 
                      scale(n_crops) * 
                      scale(prop_cropland) * 
                      scale(bin_ag) + 
                      (1|species) +
                      (1|state) + 
                      (1|county),
                    data = model.vars.df,
                    weights = bumble_abun,
                    family = binomial(link = "logit"),
                    control = glmerControl(optimizer = "bobyqa",
                                 optCtrl = list(maxfun = 2e5)))
summary(global.glm)

interact_plot(global.glm,
              # pred = bin_ag,
              pred = prop_cropland,
              modx = n_crops,
              # mod2 = prop_cropland,
              mod2 = bin_ag,
              data = model.vars.df,
              mod2.values = c(1870, 1900, 1950, 1982, 2017),
              # mod2.values = c(0.1, 0.5, 0.9),
              outcome.scale = "response",
              interval = TRUE,
              x.label = "Year",
              y.label = "Probability of occurrence")


interact_plot(global.glm,
              pred = bin_ag,
              modx = n_crops,
              mod2 = prop_cropland,
              data = model.vars.df,
              mod2.values = c(0.1, 0.5, 0.9),
              outcome.scale = "response",
              interval = TRUE,
              x.label = "Year",
              y.label = "Probability of occurrence")
                      
```

